
\section{Stochastic Gradient Descent\label{sec:Stochastic-Gradient-Descent}}

In this section, we study problems of the form $\sum f_{i}$ where
each $f_{i}$ is convex and the sum can be either infinite or finite.

\subsection{Stochastic Problem}

In this part, we discuss the problem
\[
\min_{x}F(x),\mbox{ where }F(x)\defeq\E_{f\sim\mathcal{D}}f(x)
\]
where $\mathcal{D}$ is a distribution of convex functions in $\R^{d}$.
The goal is to find the minimizer $x^{*}=\argmin_{x}F(x)$. Suppose
we observed samples $f_{1},f_{2},\cdots,f_{T}$ from $\mathcal{D}$.
Ideally, we wish to approximate $x^{*}$ by the empirical risk minimizer
\[
x^{(T)}_{\mathrm{ERM}}=\arg\min_{x}F_{T}(x)\defeq\frac{1}{T}\sum^{T}_{i=1}f_{i}(x).
\]
It is known that $x^{(T)}_{\mathrm{ERM}}$ is optimal in a certain
sense in spite of its computational cost. Therefore, to discuss the
efficiency of an optimization algorithm for $F(x)$, it is helpful
to consider the ratio
\[
\frac{\E F(x^{(T)})-F(x^{*})}{\E F(x^{(T)}_{\mathrm{ERM}})-F(x^{*})}.
\]
We will first discuss the term $\E F(x^{(T)}_{\mathrm{ERM}})-F(x^{*})$.
As an example, consider the simplest one-dimensional problem $F(x)=\E_{b\sim N(0,1)}(x-b)^{2}$.
Note that $x^{(T)}_{\mathrm{ERM}}$ is simply the average of $T$
standard normal variables and hence 
\begin{align*}
\E F(x^{(T)}_{\mathrm{ERM}})-F(x^{*}) & =\E_{b_{1},b_{2},\cdots,b_{T}}\E_{b}(x^{(T)}_{\mathrm{ERM}}-b)^{2}-b^{2}\\
 & =\E_{b_{1},b_{2},\cdots,b_{T}}(x^{(T)}_{\mathrm{ERM}})^{2}\\
 & =\E_{b_{1},b_{2},\cdots,b_{T}}(\frac{1}{T}\sum^{T}_{i=1}b_{i})^{2}=\frac{1}{T}.
\end{align*}
In general, the following lemma shows that
\[
\E F(x^{(T)}_{\mathrm{ERM}})-F(x^{*})\rightarrow\frac{\sigma^{2}}{T}\quad\text{where }\sigma^{2}=\frac{1}{2}\E\|\nabla f(x^{*})\|^{2}_{(\nabla^{2}F(x^{*}))^{-1}}.
\]
Here,$\frac{\sigma^{2}}{T}$ is called the Cramer-Rao bound. 

\begin{lem}
Suppose that $f$ is $\mu$-strongly convex with Lipschitz Hessian
for all $f\sim\mathcal{D}$ for some $\mu>0$. Suppose that $\E_{f\sim\mathcal{D}}\|\nabla f(x^{*})\|^{2}<+\infty$.
Then, we have that
\[
\lim_{N\rightarrow+\infty}\frac{\E F(x^{(N)}_{\mathrm{ERM}})-F(x^{*})}{\sigma^{2}/N}=1.
\]
\end{lem}

\begin{rem*}
The statement holds with weaker assumptions and the rate of convergence
can be made quantitative.
\end{rem*}
\begin{proof}
We first prove $x^{(N)}_{\mathrm{ERM}}\rightarrow x^{*}$ as $N\rightarrow+\infty$.
By the optimality condition of $x^{(N)}_{\mathrm{ERM}}$, and using
Taylor's theorem, for some $\widetilde{x}$, we have that
\begin{equation}
0=\nabla F_{N}(x^{(N)}_{\mathrm{ERM}})=\nabla F_{N}(x^{*})+\nabla^{2}F_{N}(\widetilde{x})(x^{(N)}_{\mathrm{ERM}}-x^{*}).\label{eq:error_F_N}
\end{equation}
By the $\mu$-strong convexity, we have
\[
\|x^{(N)}_{\mathrm{ERM}}-x^{*}\|^{2}_{2}\leq\frac{1}{\mu^{2}}\|\nabla F_{N}(x^{*})\|^{2}.
\]
Since $\E_{f\sim\mathcal{D}}\nabla f(x^{*})=\nabla F(x^{*})=0$, we
have
\begin{align*}
\E\|x^{(N)}_{\mathrm{ERM}}-x^{*}\|^{2}_{2} & \leq\frac{1}{\mu^{2}}\E\frac{1}{N^{2}}\sum_{i}\|\nabla f_{i}(x^{*})\|^{2}\\
 & =\frac{1}{\mu^{2}N}\E_{f\sim\mathcal{D}}\|\nabla f(x^{*})\|^{2}
\end{align*}
Therefore, $x^{(N)}_{\mathrm{ERM}}\rightarrow x^{*}$ as $N\rightarrow+\infty$.

Now, to compute the error, Taylor expansion of $F$ at $x^{*}$ shows
that
\[
F(x^{(N)}_{\mathrm{ERM}})-F(x^{*})=\frac{1}{2}(x^{(N)}_{\mathrm{ERM}}-x^{*})^{\top}\nabla^{2}F(\overline{x})(x^{(N)}_{\mathrm{ERM}}-x^{*})
\]
for some $\overline{x}$ between $x^{*}$ and $x^{(N)}_{\mathrm{ERM}}$.
Using this and (\ref{eq:error_F_N}) gives
\[
F(x^{(N)}_{\mathrm{ERM}})-F(x^{*})=\frac{1}{2}\nabla F_{N}(x^{*})^{\top}(\nabla^{2}F_{N}(\widetilde{x}))^{-1}\nabla^{2}F(\overline{x})(\nabla^{2}F_{N}(\widetilde{x}))^{-1}\nabla F_{N}(x^{*}).
\]

Since $x^{(N)}_{\mathrm{ERM}}\rightarrow x^{*}$ and $\nabla^{2}F_{N}\succeq\mu I$,
we have $(\nabla^{2}F_{N}(\widetilde{x}))^{-1}\nabla^{2}F(\overline{x})(\nabla^{2}F_{N}(\widetilde{x}))^{-1}\rightarrow(\nabla^{2}F(x^{*}))^{-1}$.
Hence, we have
\begin{align*}
\lim_{N\rightarrow\infty}\E N\cdot(F(x^{(N)}_{\mathrm{ERM}})-F(x^{*})) & =\lim_{N\rightarrow\infty}\frac{N}{2}\E\nabla F_{N}(x^{*})^{\top}(\nabla^{2}F(x^{*}))^{-1}\nabla F_{N}(x^{*})\\
 & =\frac{1}{2}\E\|\nabla f(x^{*})\|^{2}_{(\nabla^{2}F(x^{*}))^{-1}}.
\end{align*}
\end{proof}
Now, we discuss how to achieve a bound similar to $\sigma^{2}/T$
using stochastic gradient descent. Since gradient descent is a first-order
method, we can only achieve a bound related to the $\ell_{2}$ norm,
$\E\|\nabla f(x^{*})\|^{2}_{2}$, instead of the inverse Hessian norm. 
\begin{lem}
\label{lem:change_variance}Suppose $f$ has $L$-Lipschitz gradient
for all $f\in\mathcal{D}$. Let $x^{*}$ be the minimizer of $F$.
Then, we have
\[
\E_{f\sim\mathcal{D}}\|\nabla f(x)-\nabla f(x^{*})\|^{2}_{2}\leq2L\cdot(F(x)-F(x^{*})).
\]
\end{lem}

\begin{proof}
Let $g(x)=f(x)-\nabla f(x^{*})^{\top}(x-x^{*})$. By construction,
$x^{*}$ is the minimizer of $g$ and $f(x^{*})=g(x^{*})$. Hence,
by the progress of gradient descent, we know
\[
g(x^{*})\leq g(x)-\frac{1}{2L}\|\nabla g(x)\|^{2}.
\]
Rearranging the term, we have
\[
\|\nabla g(x)\|^{2}\leq2L(g(x)-g(x^{*}))
\]
and hence
\[
\|\nabla f(x)-\nabla f(x^{*})\|^{2}\leq2L\cdot(f(x)-f(x^{*})-\nabla f(x^{*})^{\top}(x-x^{*})).
\]
Taking expectation, we have the result.
\end{proof}
\begin{algorithm2e}[H]

\caption{$\mathtt{StochasticGradientDescent}$ (SGD)}

\SetAlgoLined

\textbf{Input:} Initial point $x^{(0)}\in\R^{d}$, step size $h>0$.
Access to $\mathcal{D}$, a distribution on convex functions.

\For{$k=0,1,\cdots,T$}{

Sample $f\sim\mathcal{D}$.

$x^{(k+1)}\leftarrow x^{(k)}-h\cdot\nabla f(x^{(k)})$.

}

\end{algorithm2e}
\begin{thm}
\label{thm:sgd_convex_strongly}Suppose $f$ has $L$-Lipschitz gradient
for all $f\in\mathcal{D}$ and $F$ is $\mu$ strongly convex. Let
$x^{*}$ be the minimizer of $F$ and $\sigma^{2}=\frac{1}{2}\E\|\nabla f(x^{*})\|^{2}$.
Assume $\sigma^{2}<\infty$ and that each sampled gradient is independent
of the past, so $\E[\nabla f(x)\mid x]=\nabla F(x)$. For step size
$h\leq\frac{1}{4L}$, the sequence $x^{(k)}$ in $\mathtt{StochasticGradientDescent}$
satisfies
\[
\E\|x^{(k)}-x^{*}\|^{2}\leq\frac{8h\sigma^{2}}{\mu}+(1-\frac{h\mu}{2})^{k}\cdot\|x^{(0)}-x^{*}\|^{2}
\]
and
\[
\frac{1}{T}\sum^{T-1}_{k=0}\E F(x^{(k)})-F(x^{*})\leq4h\sigma^{2}+\frac{\|x^{(0)}-x^{*}\|^{2}}{hT}.
\]
\end{thm}

\begin{proof}
Note that
\[
\|x^{(k+1)}-x^{*}\|^{2}=\|x^{(k)}-x^{*}\|^{2}-2h\left\langle x^{(k)}-x^{*},\nabla f(x^{(k)})\right\rangle +h^{2}\|\nabla f(x^{(k)})\|^{2}.
\]
Conditioning on $x^{(k)}$ and taking expectation over the randomness
in $f$ and using that 
\begin{align*}
\E[\|\nabla f(x^{(k)})\|^{2}\mid x^{(k)}] & \leq2\E[\|\nabla f(x^{(k)})-\nabla f(x^{*})\|^{2}\mid x^{(k)}]+2\E\|\nabla f(x^{*})\|^{2}\\
 & \leq4L\cdot(F(x^{(k)})-F(x^{*}))+4\sigma^{2},
\end{align*}
we have
\begin{align}
\E[\|x^{(k+1)}-x^{*}\|^{2}\mid x^{(k)}] & =\|x^{(k)}-x^{*}\|^{2}-2h\left\langle x^{(k)}-x^{*},\nabla F(x^{(k)})\right\rangle +4h^{2}\sigma^{2}+4Lh^{2}\cdot(F(x^{(k)})-F(x^{*}))\nonumber \\
 & \leq\|x^{(k)}-x^{*}\|^{2}-(2h-4Lh^{2})(F(x^{(k)})-F(x^{*}))+4h^{2}\sigma^{2}.\label{eq:SGD_distance}
\end{align}
Using $h\leq\frac{1}{4L}$ and$F(x^{(k)})-F(x^{*})\geq\frac{\mu}{2}\|x^{(k)}-x^{*}\|^{2}_{2}$,
we have
\[
\E[\|x^{(k+1)}-x^{*}\|^{2}\mid x^{(k)}]\leq4h^{2}\sigma^{2}+(1-\frac{h\mu}{2})\cdot\|x^{(k)}-x^{*}\|^{2}.
\]
Taking total expectation gives the same inequality with unconditional
expectations. 

The first conclusion follows.

For the second conclusion, (\ref{eq:SGD_distance}) shows that
\[
\E F(x^{(k)})-F(x^{*})\leq\frac{\E\|x^{(k)}-x^{*}\|^{2}-\E\|x^{(k+1)}-x^{*}\|^{2}}{h}+4h\sigma^{2}.
\]
Hence, we have
\begin{align*}
\frac{1}{T}\sum^{T-1}_{k=0}\E F(x^{(k)})-F(x^{*}) & \leq4h\sigma^{2}+\frac{\|x^{(0)}-x^{*}\|^{2}}{hT}
\end{align*}
\end{proof}
\begin{xca}
Apply the theorem above twice to show that one can achieve error $\tilde{O}(\frac{\sigma^{2}}{\mu T})$,
which is roughly the same as the Cramer-Rao bound. 
\end{xca}

We note that for many algorithms in this book, such as mirror descent,
the stochastic version is obtained by replacing the gradient $\nabla F$
by the gradient of a sample, $\nabla f$.

\subsection{Finite Sum Problem}

When $\nabla f(x^{*})=0$ for all $f\sim\mathcal{D}$, stochastic
gradient descent converges exponentially. When the function is given
by a finite sum $F=\frac{1}{n}\sum f_{i}$, we can reduce the variance
at $x^{*}$ by replacing 
\[
\nabla\widetilde{f}_{i}(x)=\nabla f_{i}(x)-\nabla f_{i}(x^{(0)})+\nabla F(x^{(0)}).
\]
Note that if $x^{(0)}$ is close to $x^{*}$, then $\nabla\widetilde{f}_{i}(x^{*})$
is small because both $\nabla f_{i}(x^{*})-\nabla f_{i}(x^{(0)})$
and $\nabla F(x^{(0)})$ are small. Formally, the variance for $\widetilde{f}$
is bounded as follows:
\begin{lem}
\label{lem:variance_reduction}Suppose $f$ has $L$-Lipschitz gradient
for all $f\in\mathcal{D}$. For any $f\in\mathcal{D}$, let $\nabla\widetilde{f}(x)=\nabla f(x)-\nabla f(x^{(0)})+\nabla F(x^{(0)})$
for some fixed $x^{(0)}$. Then, we have
\[
\E\|\nabla\widetilde{f}(x^{*})\|^{2}\leq8L\cdot(F(x^{(0)})-F(x^{*})).
\]
\end{lem}

\begin{proof}
Note that
\begin{align*}
\E\|\nabla\widetilde{f}(x^{*})\|^{2} & \leq2\E\|\nabla f(x^{*})-\nabla f(x^{(0)})\|^{2}+2\|\nabla F(x^{(0)})-\nabla F(x^{*})\|^{2}\\
 & \leq2\E\|\nabla f(x^{*})-\nabla f(x^{(0)})\|^{2}+2\E\|\nabla f(x^{(0)})-\nabla f(x^{*})\|^{2}\\
 & \leq4L(F(x^{(0)})-F(x^{*}))+4L(F(x^{(0)})-F(x^{*}))
\end{align*}
where we used Lemma \ref{lem:change_variance} at the end.
\end{proof}
\begin{algorithm2e}[H]

\caption{$\mathtt{StochasticVarianceReducedGradient}$ (SVRG)}

\SetAlgoLined

\textbf{Input:} Initial point $x^{(0)}\in\R^{d}$, step size $h>0$,
restart time $m$.

$\widetilde{x}=x^{(0)}$.

\For{$k=0,1,\cdots,T$}{

\If{$k$ is a multiple of $m$ and $k\neq0$}{

$x^{(k)}=\frac{1}{m}\sum^{m}_{j=1}x^{(k-j)}$.

$\widetilde{x}=x^{(k)}$.

}

Sample $i\in[n]$ uniformly.

$x^{(k+1)}\leftarrow x^{(k)}-h\cdot(\nabla f_{i}(x^{(k)})-\nabla f_{i}(\widetilde{x})+\nabla F(\widetilde{x}))$.

}

\end{algorithm2e}
\begin{thm}
\label{thm:stochastic_method}Suppose each$f_{i}$ has $L$-Lipschitz
gradient and$F$ is $\mu$ strongly convex. Let $x^{*}$ be the minimizer
of $F$. Here $F=\frac{1}{n}\sum^{n}_{i=1}f_{i}$, and the index $i$
in each inner step is sampled uniformly from $[n]$. For step size
$h=\frac{1}{64L}$ and $m=\frac{256L}{\mu}$, the sequence $x^{(km)}$
in $\mathtt{StochasticVarianceReducedGradient}$ satisfies
\[
\E F(x^{(km)})-F(x^{*})\leq\frac{1}{2^{k}}\cdot(F(x^{(0)})-F(x^{*})).
\]
In particular, it takes $O((n+\frac{L}{\mu})\log(\frac{1}{\epsilon}))$
gradient computations to find $x$ such that $\E F(x)-F(x^{*})\leq\epsilon\cdot(F(x^{(0)})-F(x^{*}))$.
\end{thm}

\begin{rem*}
Gradient descent gives $O(n\frac{L}{\mu}\log(\frac{1}{\epsilon}))$
instead because computing $\nabla F$ involves computing $n$ gradients
$\nabla f$.
\end{rem*}
\begin{proof}
The algorithm consists of $\frac{T}{m}$ phases. For the first phase,
we compute $\nabla F(x^{(0)})$. Lemma \ref{lem:variance_reduction}
shows that 
\[
\E\|\nabla f(x^{*})-\nabla f(x^{(0)})+\nabla F(x^{(0)})\|^{2}\leq8L\cdot(F(x^{(0)})-F(x^{*})).
\]
Hence, Theorem \ref{thm:sgd_convex_strongly} shows that
\begin{align*}
\frac{1}{m}\sum^{m-1}_{k=0}\E F(x^{(k)})-F(x^{*}) & \leq16hL\cdot(F(x^{(0)})-F(x^{*}))+\frac{F(x^{(0)})-F(x^{*})}{\mu hm}\\
 & \leq\frac{1}{2}(F(x^{(0)})-F(x^{*})).
\end{align*}
Hence, the error decreases by half each phase. This shows the result.
\end{proof}

\section{Coordinate Descent}

If some of the coordinates are more important than others for optimization,
it makes sense to update the important coordinates more often.
\begin{lem}
\label{lem:coordinate_descent}Given a convex function $f$, suppose
that $\frac{\partial^{2}}{\partial x^{2}_{i}}f(x)\leq L_{i}$ for
all $x$ and let $L=\sum L_{i}$. If we sample coordinate $i$ with
probability $p_{i}=\frac{L_{i}}{L}$, then, 
\[
\E_{i}f(x-\frac{1}{L_{i}}\frac{\partial}{\partial x_{i}}f(x)e_{i})\leq f(x)-\frac{1}{2L}\norm{\nabla f(x)}^{2}_{2}.
\]
\end{lem}

\begin{proof}
Note that the function $\zeta(t)=f(x+te_{i})$ is $L_{i}$ smooth.
Hence, we have that
\[
f(x-\frac{1}{L_{i}}\frac{\partial}{\partial x_{i}}f(x)e_{i})\leq f(x)-\frac{1}{2L_{i}}\left(\frac{\partial f}{\partial x_{i}}(x)\right)^{2}.
\]
Since we sample coordinate $i$ with probability $p_{i}=\frac{L_{i}}{L}$,
we have that
\begin{align*}
\E f(x-\frac{1}{L_{i}}\frac{\partial}{\partial x_{i}}f(x)e_{i}) & \le f(x)-\sum_{i}\frac{L_{i}}{L}\frac{1}{2L_{i}}\left(\frac{\partial f}{\partial x_{i}}(x)\right)^{2}\\
 & =f(x)-\frac{1}{2L}\sum_{i}\left(\frac{\partial f}{\partial x_{i}}(x)\right)^{2}\\
 & =f(x)-\frac{1}{2L}\norm{\nabla f(x)}^{2}_{2}.
\end{align*}
\end{proof}
By the same proof as Theorem \ref{thm:gd_convex}, we have the following:

\begin{algorithm2e}[H]

\caption{$\mathtt{CoordinateDescent}$ (CD)}

\SetAlgoLined

\textbf{Input:} Initial point $x^{(0)}\in\R^{d}$, step size $h>0$.

\For{$k=0,1,\cdots,T$}{

Sample $i$ with probability proportional to $L_{i}$.

$x^{(k+1)}\leftarrow x^{(k)}-\frac{1}{L_{i}}\frac{\partial}{\partial x_{i}}f(x^{(k)})e_{i}$.

}

\end{algorithm2e}
\begin{thm}[Coordinate Descent Convergence]
\label{thm:coordinate_descent}Given a convex function $f$, suppose
that $\frac{\partial^{2}}{\partial x^{2}_{i}}f(x)\leq L_{i}$ for
all $x$ and let $L=\sum L_{i}$. Consider the algorithm $x^{(k+1)}\leftarrow x^{(k)}-\frac{1}{L_{i}}\frac{\partial}{\partial x_{i}}f(x^{(k)})e_{i}$.
Then, we have that 
\[
\E f(x^{(k)})-f(x^{*})\leq\frac{2LR^{2}}{k+4}\quad\text{with}\quad R=\max_{f(x)\leq f(x^{(0)})}\norm{x-x^{*}}_{2}
\]
for any minimizer $x^{*}$ of $f$. If $f$ is $\mu$ strongly convex,
then we also have 
\[
\E f(x^{(k)})-f(x^{*})\leq(1-\frac{\mu}{L})^{\Omega(k)}(f(x^{(0)})-f(x^{*})).
\]
\end{thm}

\begin{rem*}
Note that $\frac{L}{d}\leq\Lip(\nabla f)\leq L$. Therefore, gradient
descent takes at least $\frac{1}{d}$ times as many steps as coordinate
descent while each step takes $d$ times longer (usually).
\end{rem*}

