
\section{Interior Point Method for Linear Programs\label{sec:IPM}}

In this section, we study interior point methods. In practice, this
method reduces the problem of optimizing a convex function to solving
a small number (``typically less than $30$'') linear systems. In
theory, it reduces the problem to $\tilde{O}(\sqrt{n})$ linear systems.

We start by describing interior point method for linear programs.
We establish a polynomial bound and then later discuss implementation
details.

\subsection{Basic Properties}

We first consider the primal problem
\[
(\mathrm{P}):\quad\quad\min_{x}c^{\top}x\text{ subject to }Ax=b,x\geq0
\]
where $A\in\R^{m\times n}$. The difficulty of linear programs is
the constraint $x\geq0$. Without this constraint, we can simply solve
it as a linear system. One natural idea to solve linear programs is
to replace the hard constraint $x\geq0$ by some smooth function.
So, let us consider the following ``regularized'' version of the
linear program for some $t\ge0$:
\[
(\mathrm{P}_{t}):\quad\quad\min_{x}c^{\top}x-t\sum_{i=1}^{n}\ln x_{i}\text{ subject to }Ax=b.
\]
We will explain the reason of choosing $\ln x$ in more detail later.
For now, we can think it as a nice function that blows up to $\infty$
as $x$ approaches zero.

One can think that $-\ln x$ gives a force from every constraint $x\geq0$
to make sure $x\geq0$ is true. Since the gradient of $-\ln x$ blows
up when $x=0$, when $x$ is close enough, the force is large enough
to counter the cost $c$. When $t\rightarrow0$, then the problem
$(\mathrm{P}_{t})$ is closer to the original problem $(\mathrm{P})$
and hence the minimizer of $(\mathrm{P}_{t})$ is closer to a minimizer
of $(\mathrm{P})$.

First, we give a formula for the minimizer of $(\mathrm{P}_{t})$.
\begin{lem}[Existence and Uniqueness of central path]
If the polytope $\{Ax=b,x\geq0\}$ has an interior, then the optimum
of $(\mathrm{P}_{t})$ is unique and is given by the solution of the
following system:
\begin{align*}
xs & =t,\\
Ax & =b,\\
A^{\top}y+s & =c,\\
(x,s) & \geq0
\end{align*}
where the variables $s_{i}$ are additional ``slack'' variables,
and $xs=t$ is shorthand of $x_{i}s_{i}=t$ for all $i$.
\end{lem}

\begin{proof}
The optimality condition, using dual variables $y$(Lagrangian multipliers)
for the Lagrangian of $Ax=b$ is given by
\[
c-\frac{t}{x}=A^{\top}y.
\]
Write $s_{i}=\frac{t}{x_{i}}$, to get the formula. The solution is
unique because the function $-\ln x$ is strictly convex.
\end{proof}
\begin{defn}
We define the central path ${\cal C}_{t}=(x^{(t)},y^{(t)},s^{(t)})$
as the sequence of points satisfying
\begin{align*}
x^{(t)}s^{(t)} & =t,\\
Ax^{(t)} & =b,\\
A^{\top}y^{(t)}+s^{(t)} & =c,\\
(x^{(t)},s^{(t)}) & \geq0.
\end{align*}
\end{defn}

To give another interpretation of the central path, note that the
dual problem is
\[
(\mathrm{D}):\quad\quad\max_{y,s}b^{\top}y\text{ subject to }A^{\top}y+s=c,s\geq0.
\]
Note that for any feasible $x$ and $y$, we have that
\[
0\leq c^{\top}x-b^{\top}y=c^{\top}x-x^{\top}A^{\top}y=x^{\top}s.
\]
Hence, $(x,y,s)$ solves the linear program if it satisfies the central
path equation with $t=0$. Therefore, following the central path is
a balanced way to decrease $x_{i}s_{i}$ uniformly to $0$. We can
formalize the intuition that for small $t$, $x^{(t)}$ is a good
approximation of the primal solution. In fact $t$ itself is a bound
on the error of the current solution.
\begin{lem}[Duality Gap]
We have that
\[
\text{Duality Gap}=c^{\top}x^{(t)}-b^{\top}y^{(t)}=c^{\top}x^{(t)}-\left(x^{(t)}\right)^{\top}A^{\top}y^{(t)}=\left(x^{(t)}\right)^{\top}s^{(t)}=tn.
\]
\end{lem}

The interior point method follows the following framework:
\begin{enumerate}
\item Find $\mathcal{C}_{1}$
\item Until $t<\frac{\varepsilon}{n}$,
\begin{enumerate}
\item Use $\mathcal{C}_{t}$ to find $\mathcal{C}_{(1-h)t}$ for $h=\frac{1}{10\sqrt{n}}$.
\end{enumerate}
\end{enumerate}
Note that this algorithm only finds a solution with $\varepsilon$
error. If the linear program is integral, we can simply stop at small
enough $\varepsilon$ and round it off to the closest integral point.

To see why there is a pair $(x,s)\in\PriR^{\circ}\times\DualR^{\circ}$
satisfying the invariant $xs=t$, we prove the following generalization.
\begin{lem}[Quadrant Representation of Primal-Dual]
\label{lem:primal_dual_representation}Suppose $\PriR$ is non-empty
and bounded. For any positive vector $\mu\in\R_{>0}^{n}$, there is
an unique pair $(x_{\mu},s_{\mu})\in\PriR^{\circ}\times\DualR^{\circ}$
such that $x_{\mu}s_{\mu}=\mu$. Furthermore, $x_{\mu}=\arg\min_{x\in\PriR}f_{\mu}(x)$
where
\[
f_{\mu}(x)=c^{\top}x-\sum_{i=1}^{n}\mu_{i}\ln x_{i}.
\]
\end{lem}

\begin{proof}
Fix $\mu\in\R_{>0}^{n}$. We define $x_{\mu}=\arg\min_{x\in\PriR}f_{\mu}(x)$
and prove that $(x_{\mu},s_{\mu})\in\PriR^{\circ}\times\DualR^{\circ}$
with $x_{\mu}s_{\mu}=\mu$ for some $s_{\mu}$. Since $\PriR$ is
non-empty and bounded and since $f_{\mu}$ is strictly convex, such
an $x_{\mu}$ exists. Furthermore, since $f_{\mu}(x)\rightarrow+\infty$
as $x_{i}\rightarrow0$ for any $i$, we have that $x_{\mu}\in\PriR^{\circ}$.

By the KKT optimality condition for $f_{\mu}$, there is a vector
$y$ such that
\[
\nabla f_{\mu}(x_{\mu})=c-\frac{\mu}{x_{\mu}}=\ma^{\top}y.
\]
Define $s_{\mu}=\frac{\mu}{x_{\mu}}$, then one can check that $s_{\mu}\in\DualR^{\circ}$
and $x_{\mu}s_{\mu}=\mu$.
\end{proof}

\subsection{Following the central path}

During the algorithm, we maintain a point $(x,y,s)$ such that $Ax=b$,
$A^{\top}y+s=c$ and $x_{i}s_{i}$ is close to $t$ for all $i$.
We show how to find a feasible $(x+\delta_{x},y+\delta_{y},s+\delta_{s})$
such that it is even closer to $t$. We can write the equation as
follows:
\begin{align*}
(x+\delta_{x})(s+\delta_{s}) & \thickapprox t,\\
A(x+\delta_{x}) & =b,\\
A^{\top}(y+\delta_{y})+(s+\delta_{s}) & =c.
\end{align*}
(Omitted the non-negative conditions.) Using our assumption on $(x,y,s)$
and noting that $\delta_{x}\cdot\delta_{s}$ is small, the equation
can simplified as follows. We use the notation $X=\Diag(x)$, $S=\Diag(s)$.
\[
\left[\begin{array}{ccc}
0 & A^{\top} & I\\
A & 0 & 0\\
S & 0 & X
\end{array}\right]\left[\begin{array}{c}
\delta_{x}\\
\delta_{y}\\
\delta_{s}
\end{array}\right]=\left[\begin{array}{c}
0\\
0\\
t-xs
\end{array}\right].
\]
This is a linear system and hence we can solve it exactly.
\begin{xca}
\label{exer:d_central}Let $r=t-xs$. Prove that $S\delta_{x}=(I-\overline{P})r$
and $X\delta_{s}=\overline{P}r$ where $\overline{P}=XA^{\top}(AS^{-1}XA^{\top})^{-1}AS^{-1}$.

First, we show that $x^{\new}=x+\delta_{x}$ and $s^{\new}=s+\delta_{s}$
are feasible.
\end{xca}

\begin{lem}
Suppose $\sum_{i}\left(x_{i}s{}_{i}-t\right)^{2}\leq\varepsilon^{2}t^{2}$
with $\varepsilon<\frac{1}{2}$. Then, $x_{i}^{\new}>0$ and $s_{i}^{\new}>0$
for all $i$.
\end{lem}

\begin{proof}
Note that $\overline{P}^{2}=\overline{P}$, where $\overline{P}$
is as defined in the exercise above. However, in general $\overline{P}\neq\overline{P}^{T},$
i.e., $\overline{P}$ might not be an orthogonal projection matrix.
It will be convenient to consider the orthogonal projection matrix
$P=S^{-\frac{1}{2}}X^{\frac{1}{2}}A^{\top}(AS^{-1}XA^{\top})^{-1}AS^{-\frac{1}{2}}X^{\frac{1}{2}}$.
Note that
\[
X^{-1}\delta_{x}=S^{-\frac{1}{2}}X^{-\frac{1}{2}}(I-P)S^{-\frac{1}{2}}X^{-\frac{1}{2}}r.
\]
By the assumption for each $i$, $x_{i}s_{i}\ge(1-\varepsilon)t$.
Therefore, we have
\begin{align*}
\|X^{-1}\delta_{x}\|_{2} & \leq\frac{1}{\sqrt{(1-\varepsilon)t}}\|(I-P)S^{-\frac{1}{2}}X^{-\frac{1}{2}}r\|_{2}\\
 & \leq\frac{1}{\sqrt{(1-\varepsilon)t}}\|S^{-\frac{1}{2}}X^{-\frac{1}{2}}r\|_{2}\\
 & \leq\frac{1}{(1-\varepsilon)t}\|r\|_{2}\leq\frac{\varepsilon}{1-\varepsilon}.
\end{align*}
Similarly, we have $S^{-1}\delta_{s}=S^{-\frac{1}{2}}X^{-\frac{1}{2}}PS^{-\frac{1}{2}}X^{-\frac{1}{2}}r$.
Hence, we have $\|S^{-1}\delta_{s}\|_{2}\leq\frac{\varepsilon}{1-\varepsilon}$.
Therefore, when $\epsilon<\frac{1}{2}$, we have both $\|X^{-1}\delta_{x}\|_{\infty}$
and $\|S^{-1}\delta_{s}\|_{\infty}$ less than $1$, which shows that
both $x^{\new}$ and $s^{\new}$ are positive. 

Next, we show that $xs$ is closer to $t$ after one Newton step.
\end{proof}
\begin{lem}
\label{lem:IPM_progress} If $\sum_{i}\left(x_{i}s{}_{i}-t\right)^{2}\leq\varepsilon^{2}t^{2}$
with $\varepsilon<\frac{1}{4}$, we have that 
\[
\sum_{i}\left(x_{i}^{\new}s_{i}^{\new}-t\right)^{2}\leq\left(\varepsilon^{4}+16\varepsilon^{5}\right)t^{2}.
\]
\end{lem}

\begin{proof}
We have that $x_{i}\delta_{s,i}+s_{i}\delta_{x,i}=t-x_{i}s_{i}$.
Using this, 
\[
\text{LHS}=\sum_{i}\left(x_{i}^{\new}s_{i}^{\new}-t\right)^{2}=\sum_{i}(x_{i}s_{i}+x_{i}\delta_{s,i}+s_{i}\delta_{x,i}+\delta_{x,i}\delta_{s,i}-t)^{2}=\sum_{i}\delta_{x,i}^{2}\delta_{s,i}^{2}\leq\left((1+\epsilon)t\right)^{2}\cdot\sum_{i}\left(\frac{\delta_{x,i}}{x_{i}}\right)^{2}\left(\frac{\delta_{s,i}}{s_{i}}\right)^{2}
\]
where in the last step we used $x_{i}^{2}s_{i}^{2}\leq(1+\varepsilon)^{2}t^{2}$.
Using the previous lemma, we have that
\begin{align*}
\text{LHS} & \leq\left((1+\varepsilon)t\right)^{2}\cdot\|X^{-1}\delta_{x}\|_{4}^{2}\|S^{-1}\delta_{s}\|_{4}^{2}\\
 & \leq\left((1+\varepsilon)t\right)^{2}\cdot\|X^{-1}\delta_{x}\|_{2}^{2}\|S^{-1}\delta_{s}\|_{2}^{2}\\
 & \leq\left((1+\varepsilon)t\right)^{2}\left(\frac{\varepsilon}{1-\varepsilon}\right)^{4}\\
 & \leq\left(\varepsilon^{4}+16\varepsilon^{5}\right)t^{2}
\end{align*}
Using this, we have the main theorem. 
\end{proof}
\begin{thm}
We can solve a linear program to within $\delta$ ``error'' (see
Lemma \ref{lem:feasible_LP}) in $O(\sqrt{n}\log(\frac{1}{\delta}))$
iterations and each iteration only needs to solve a linear system.
\end{thm}

\begin{proof}
Let $\Phi=\sum_{i}\left(x_{i}s{}_{i}-t\right)^{2}$ be the error of
the current iteration. We always maintain $\Phi\leq\frac{t^{2}}{16}$
for the current $(x,y,s)$ and $t$. At each step, we use Lemma \ref{lem:IPM_progress}
which makes $\Phi\leq\frac{t^{2}}{50}$. Then, we decrease $t$ to
$t(1-\frac{1}{10\sqrt{n}})$. Note that
\[
\sum_{i}\left(x_{i}s{}_{i}-t(1-h)\right)^{2}\leq2\Phi+2t^{2}h^{2}n\leq\frac{2t^{2}}{50}+\frac{2t^{2}}{100}\leq\frac{(t(1-h))^{2}}{16}.
\]
Therefore, the invariant is preserved after each step. Since $t$
is decreased by a $\left(1-\frac{1}{10\sqrt{n}}\right)$ factor each
step, it takes $O(\sqrt{n}\log(\frac{1}{\delta}))$ to decrease $t$
from $1$ to $\delta^{2}$.
\end{proof}


\subsection{Finding the initial point}

The first question is to find $\mathcal{C}_{1}$. This can be handled
by extending the problem to slightly higher dimension. To the reader
familiar with the Simplex method, this might be reminiscent of the
two phases of the simplex method, where the purpose of the first phase
is to find a feasible initial solution.The idea of the construction
is simple: if we simply solve $Ax=b$, the solution could have negative
components. So we will allow negative components but penalize them
heavily in the objective.
\begin{lem}
\label{lem:feasible_LP} Consider a linear program $\min_{Ax=b,x\geq0}c^{\top}x$
with $n$ variables and $d$ constraints. Assume that 
\begin{enumerate}
\item Diameter: For any $x\geq0$ with $Ax=b$, we have that $\|x\|_{\infty}\leq R$.
\item Lipschitz constant of the objective: $\|c\|_{\infty}\leq L$. 
\end{enumerate}
For any $0<\delta\leq1$, the modified linear program $\min_{\overline{A}\overline{x}=\overline{b},\overline{x}\geq0}\overline{c}^{\top}\overline{x}$
with 
\[
\overline{A}=\left[\begin{array}{ccc}
A & 0 & \frac{1}{R}b-A1_{n}\\
1_{n}^{\top} & 1 & 0
\end{array}\right],\overline{b}=\left[\begin{array}{c}
\frac{1}{R}b\\
n+1
\end{array}\right]\text{, and }\overline{c}=\left[\begin{array}{c}
\delta/L\cdot c\\
0\\
1
\end{array}\right]
\]
satisfies the following:
\begin{enumerate}
\item $\overline{x}=\left[\begin{array}{c}
1_{n}\\
1\\
1
\end{array}\right]$, $\overline{y}=\left[\begin{array}{c}
0_{d}\\
-1
\end{array}\right]$ and $\overline{s}=\left[\begin{array}{c}
1_{n}+\frac{\delta}{L}\cdot c\\
1\\
1
\end{array}\right]$ are feasible primal dual vectors.
\item For any feasible primal dual vectors $(\overline{x},\overline{y},\overline{s})$
with duality gap at most $\delta^{2}$, the vector $\hat{x}=R\cdot\overline{x}_{1:n}$
($\overline{x}_{1:n}$ are the first $n$ coordinates of $\overline{x}$)
is an approximate solution to the original linear program in the following
sense 
\begin{align*}
c^{\top}\hat{x} & \leq\min_{Ax=b,x\geq0}c^{\top}x+LR\cdot\delta,\\
\|A\hat{x}-b\|_{1} & \leq4n\delta\cdot\left(R\sum_{i,j}|A_{i,j}|+\|b\|_{1}\right),\\
\hat{x} & \geq0.
\end{align*}
\end{enumerate}
\end{lem}

\textbf{Part 1.} For the first result, straightforward calculations
show that $(\overline{x},\overline{y},\overline{s})\in\R^{(n+2)\times(d+1)\times(n+2)}$
are feasible, i.e.,
\[
\overline{A}\overline{x}=\begin{bmatrix}A & 0 & \frac{1}{R}b-A1_{n}\\
1_{n}^{\top} & 1 & 0
\end{bmatrix}\cdot\begin{bmatrix}1_{n}\\
1\\
1
\end{bmatrix}=\begin{bmatrix}\frac{1}{R}b\\
n+1
\end{bmatrix}=\overline{b}
\]
and 
\begin{align*}
\overline{A}^{\top}\overline{y}+\overline{s}= & ~\begin{bmatrix}A^{\top} & 1_{n}\\
0 & 1\\
\frac{1}{R}b^{\top}-1_{n}^{\top}A^{\top} & 0
\end{bmatrix}\cdot\begin{bmatrix}0_{d}\\
-1
\end{bmatrix}+\begin{bmatrix}1_{n}+\frac{\delta}{L}\cdot c\\
1\\
1
\end{bmatrix}\\
= & ~\begin{bmatrix}-1_{n}\\
-1\\
0
\end{bmatrix}+\begin{bmatrix}1_{n}+\frac{\delta}{L}\cdot c\\
1\\
1
\end{bmatrix}\\
= & ~\begin{bmatrix}\frac{\delta}{L}\cdot c\\
0\\
1
\end{bmatrix}=\overline{c}
\end{align*}

\textbf{Part 2.} For the second result, we let 
\[
\text{OPT}=\min_{Ax=b,x\geq0}c^{\top}x,\text{~~~and,~~~}\overline{\text{OPT}}=\min_{\overline{A}\overline{x}=\overline{b},\overline{x}\geq0}\overline{c}^{\top}\overline{x}
\]
For any optimal $x\in\R^{n}$ in the original LP, we consider the
following $\overline{x}\in\R^{n+2}$ 
\begin{align}
\overline{x}=\begin{bmatrix}\frac{1}{R}x\\
n+1-\frac{1}{R}\sum_{i=1}^{n}x_{i}\\
0
\end{bmatrix}\label{eq:another_ov_x_in_appendix}
\end{align}
and $\overline{c}\in\R^{n+2}$ 
\begin{align}
\overline{c}=\begin{bmatrix}\frac{\delta}{L}\cdot c^{\top}\\
0\\
1
\end{bmatrix}\label{eq:another_ov_c_in_appendix}
\end{align}
We want to argue that $\overline{x}\in\R^{n+2}$ is feasible in the
modified LP. It is obvious that $\overline{x}\geq0$, it remains to
show $\overline{A}\overline{x}=\overline{b}\in\R^{d+1}$. We have
\begin{align*}
\overline{A}\overline{x}=\begin{bmatrix}A & 0 & \frac{1}{R}b-A1_{n}\\
1_{n}^{\top} & 1 & 0
\end{bmatrix}\cdot\begin{bmatrix}\frac{1}{R}x\\
n+1-\frac{1}{R}\sum_{i=1}^{n}x_{i}\\
0
\end{bmatrix}=\begin{bmatrix}\frac{1}{R}Ax\\
n+1
\end{bmatrix}=\begin{bmatrix}\frac{1}{R}b\\
n+1
\end{bmatrix}=\overline{b},
\end{align*}
where the third step follows from $Ax=b$, and the last step follows
from definition of $\overline{b}$.

Therefore, using the definition of $\overline{x}$ in (\ref{eq:another_ov_x_in_appendix})
we have that 
\begin{align}
\overline{\text{OPT}}\leq\overline{c}^{\top}\overline{x}=\begin{bmatrix}\frac{\delta}{L}\cdot c^{\top} & 0 & 1\end{bmatrix}\cdot\begin{bmatrix}\frac{1}{R}x\\
n+1-\frac{1}{R}\sum_{i=1}^{n}x_{i}\\
0
\end{bmatrix}=\frac{\delta}{LR}\cdot c^{\top}x=\frac{\delta}{LR}\cdot\text{OPT}.\label{eq:bounding_ov_OPT_by_OPT}
\end{align}
where the first step follows from modified program is solving a minimization
problem, the second step follows from definition of $\overline{x}\in\R^{n+2}$
(\ref{eq:another_ov_x_in_appendix}) and $\overline{c}\in\R^{n+2}$
(\ref{eq:another_ov_c_in_appendix}), the last step follows from $x\in\R^{n}$
is an optimal solution in the original linear program.

Given a feasible $(\overline{x},\overline{y},\overline{s})\in\R^{(n+2)\times(d+1)\times(n+2)}$
with duality gap $\delta^{2}$, we can write $\overline{x}=\begin{bmatrix}\overline{x}_{1:n}\\
\tau\\
\theta
\end{bmatrix}\in\R^{n+2}$ for some $\tau\geq0$, $\theta\geq0$. We can compute $\overline{c}^{\top}\overline{x}$
which is $\frac{\delta}{L}\cdot c^{\top}\overline{x}_{1:n}+\theta$.
Then, we have 
\begin{align}
\frac{\delta}{L}\cdot c^{\top}\overline{x}_{1:n}+\theta\leq\overline{\text{OPT}}+\delta^{2}\leq\frac{\delta}{LR}\cdot\overline{\text{OPT}}+\delta^{2},\label{eq:eq1_in_appendix}
\end{align}
where the first step follows from definition of duality gap, the last
step follows from (\ref{eq:bounding_ov_OPT_by_OPT}).

Hence, we can upper bound the $\overline{\text{OPT}}$ of the transformed
program as follows: 
\begin{align*}
c^{\top}\hat{x}=R\cdot c^{\top}\overline{x}_{1:n}=\frac{LR}{\delta}\cdot\frac{\delta}{L}c^{\top}\overline{x}_{1:n}\leq\frac{RL}{\delta}(\frac{\delta}{LR}\cdot\text{OPT}+\delta^{2})=\text{OPT}+LR\cdot\delta,
\end{align*}
where the first step follows by $\hat{x}=R\cdot\overline{x}_{1:n}$,
the third step follows by (\ref{eq:eq1_in_appendix}).

Note that 
\begin{align}
\frac{\delta}{L}c^{\top}\overline{x}_{1:n}\geq-\frac{\delta}{L}\|c\|_{\infty}\|\overline{x}_{1:n}\|_{1}=-\frac{\delta}{L}\|c\|_{\infty}\|\frac{1}{R}x\|_{1}\geq-\frac{\delta}{L}\|c\|_{\infty}\frac{n}{R}\|x\|_{\infty}\geq-\delta n,\label{eq:lower_bound_on_delta_L_c_ovx}
\end{align}
where the second step follows from definition $\overline{x}\in\R^{n+2}$,
and the last step follows from $\|c\|_{\infty}\leq L$ and $\|x\|_{\infty}\leq R$.

We can upper bound the $\theta$ in the following sense, 
\begin{align}
\theta\leq\frac{\delta}{LR}\cdot\overline{\text{OPT}}+\delta^{2}+\delta n\leq2n\delta+\delta^{2}\leq4n\delta\label{eq:theta_is_at_most_4n_delta}
\end{align}
where the first step follows from (\ref{eq:eq1_in_appendix}) and
(\ref{eq:lower_bound_on_delta_L_c_ovx}), the second step follows
by $\text{OPT}=\min_{Ax=b,x\geq0}c^{\top}x\leq nLR$ (because $\|c\|_{\infty}\leq L$
and $\|x\|_{\infty}\leq R$), and the last step follows from $\delta\leq1\leq n$.

The constraint in the new polytope shows that 
\begin{align*}
A\overline{x}_{1:n}+(\frac{1}{R}b-A1_{n})\theta=\frac{1}{R}b.
\end{align*}
Using $\hat{x}=Rx_{1:n}\in\R^{n}$, we have 
\begin{align*}
A\frac{1}{R}\hat{x}+(\frac{1}{R}b-A1_{n})\theta=\frac{1}{R}b.
\end{align*}
Rewriting it, we have $A\hat{x}-b=(RA1_{n}-b)\theta\in\R^{d}$ and
hence 
\begin{align*}
\|A\hat{x}-b\|_{1}=\|(RA1_{n}-b)\theta\|_{1}\leq\theta(\|RA1_{n}\|_{1}+\|b\|_{1})\leq\theta\cdot(R\|A\|_{1}+\|b\|_{1})\leq4n\delta\cdot(R\|A\|_{1}+\|b\|_{1}),
\end{align*}
where the second step follows from triangle inequality, the third
step follows from $\|A1_{n}\|_{1}\leq\|A\|_{1}$ (because the definition
of entry-wise $\ell_{1}$ norm), and the last step follows from (\ref{eq:theta_is_at_most_4n_delta}).

\subsection{Finding a feasible solution close to the Central Path}
\begin{defn}
\label{def:LP_para}We define the following parameters for the linear
program $\min_{\ma x=b,x\geq0}c^{\top}x$:
\begin{enumerate}
\item Inner radius $r$: There exists a $x\in\PriR$ such that $x_{i}\geq r$
for all $i\in[n]$.
\item Outer radius $R$: For any $x\geq0$ with $\ma x=b$, we have that
$\|x\|_{2}\leq R$.
\item Lipschitz constant $L$: $\|c\|_{2}\leq L$.
\end{enumerate}
\end{defn}

Since $\mathtt{L2Step}$ requires a feasible point near the central
path, we modify the linear program to make it happen. To satisfy the
constraint $\ma x=b$, we start the algorithm by taking a least square
solution of the constraint $\ma x=b$. Since it can be negative, we
write the variable $x=x^{+}-x^{-}$ with both $x^{+},x^{-}\geq0$.
We put a large cost vector on $x^{-}$ to ensure the solution is roughly
the same. The crux of the proof is that if we optimize this new program
well enough, we will have $x^{+}-x^{-}>0$ and hence $x^{+}-x^{-}$
gives a good starting point of the original program. We need to put
an extra constraint $1^{\top}x^{+}\leq\Lambda$ for a suitable $\Lambda$
to ensure that the problem is bounded. The precise formulation of
the modified linear program is as follows. 
\begin{defn}[Modified Linear Program]
\label{def:IPM_interior_modified}Consider a linear program $\min_{\ma x=b,x\geq0}c^{\top}x$
with inner radius $r$, outer radius $R$ and Lipschitz constant $L$.
For any $\overline{R}\geq10R$, $t\geq8L\overline{R}$, we define
the modified primal linear program by 
\[
\min_{(x^{+},x^{-},x^{\theta})\in\mathcal{P}_{\overline{R},t}}c^{\top}x^{+}+\widetilde{c}^{\top}x^{-}
\]
where 
\[
\mathcal{P}_{\overline{R},t}=\{(x^{+},x^{-},x^{\theta})\in\R_{\geq0}^{2n+1}:\ma(x^{+}-x^{-})=b,\sum_{i=1}^{n}x_{i}^{+}+x^{\theta}=\widetilde{b}\}
\]
with $x_{c}^{+}=\frac{t}{c+t/\overline{R}},x_{c}^{-}=x_{c}^{+}-\ma^{\top}(\ma\ma^{\top})^{-1}b,\widetilde{c}=t/x_{c}^{-}$,
$\widetilde{b}=\sum_{i}x_{c,i}^{+}+\overline{R}$. We define the corresponding
dual polytope by
\[
\mathcal{D}_{\overline{R},t}=\{(s^{+},s^{-},s^{\theta})\in\R_{\geq0}^{2n+1}:\ma^{\top}y+\lambda\mathbf{1}+s^{+}=c,-\ma^{\top}y+s^{-}=\widetilde{c},\lambda+s^{\theta}=0\text{ for some }y\in\R^{d}\text{ and }\lambda\in\R\}.
\]
\end{defn}

The main result about the modified program is the following.
\begin{thm}
\label{thm:IPM_interior}Given a linear program $\min_{\ma x=b,x\in\R_{\geq}^{n}}c^{\top}x$
with inner radius $r$, outer radius $R$ and Lipschitz constant $L$.
For any $0\leq\epsilon\leq\frac{1}{2}$, the modified linear program
(Definition \ref{def:IPM_interior_modified}) with $\overline{R}=\frac{5}{\epsilon}R,t=2^{16}\epsilon^{-3}n^{2}\frac{R}{r}\cdot LR$
has the following properties:
\begin{itemize}
\item The point $(x_{c}^{+},x_{c}^{-},\overline{R})$ is on the central
path of the modified program at $t$.
\item For any primal $x\defeq(x^{+},x^{-},x^{\theta})\in\mathcal{P}_{\overline{R},t}$
and dual $s\defeq(s^{+},s^{-},s^{\theta})\in\mathcal{D}_{\overline{R},t}$
such that $\frac{5}{6}LR\leq x_{i}s_{i}\leq\frac{7}{6}LR$, we have
that
\[
(x^{+}-x^{-},s^{+}-s^{\theta})\in\PriR\times\DualR
\]
and that $x_{i}^{-}\leq\epsilon x_{i}^{+}$ and $s^{\theta}\leq\epsilon s_{i}^{+}$
for all $i$.
\end{itemize}
\end{thm}

The proof of this theorem is deferred to the end of this section.
Now we state our main algorithm.

\begin{algorithm2e}[H]

\caption{$\mathtt{SlowSolveLP}(\ma,b,c,x^{(0)},\delta)$}

\label{alg:SlowSolveLP}

\textbf{Assumption: the linear program has inner radius $r$, outer
radius $R$ and Lipschitz constant $L$.}

Let $\epsilon=1/(100\sqrt{n})$, $\overline{R}=\frac{5}{\epsilon}R$,
$t=2^{16}\epsilon^{-3}n^{2}\frac{R}{r}\cdot LR$. 

\tcp{Define the modified program $\min_{\oma x=\overline{b}}\overline{c}^{\top}x$
by Definition \ref{def:IPM_interior_modified} with parameters $\overline{R}$
and $t$.}

Let $\oma=\left[\begin{array}{ccc}
\ma & -\ma & 0\\
1 & 0 & 1
\end{array}\right]$, $\overline{c}=(c,\widetilde{c})$, $\overline{b}=(b,\widetilde{b})$
where $\widetilde{c}$ and $\widetilde{b}$ are defined in Definition
\ref{def:IPM_interior_modified}.

\tcp{Write down the central path at $t$ for modified linear program.}

$\overline{x}=(x_{c}^{+},x_{c}^{-},\overline{R})$. $\overline{s}=x/t$.

$(\overline{x},\overline{s})=\mathtt{L2Step}(\oma,\overline{x},\overline{s},t,LR)$.

$(x,s)=(x^{+}-x^{-},s^{+}-s^{\theta})$ where $\overline{x}=(x^{+},x^{-},x^{\theta})$
and $\overline{s}=(s^{+},s^{-},s^{\theta})$.

$(x_{\mathrm{end}},s_{\mathrm{end}})=\mathtt{L2Step}(\ma,x^{+}-x^{-},s^{+}-s^{\theta},LR,t_{\mathrm{end}})$
with $t_{\mathrm{end}}=\delta LR/(2n)$.

\textbf{Return} $x_{\mathrm{end}}$.

\end{algorithm2e}
\begin{thm}
\label{thm:SlowSolveLP}Consider a linear program $\min_{\ma x=b,x\geq0}c^{\top}x$
with $n$ variables and $d$ constraints. Assume the linear program
has inner radius $r$, outer radius $R$ and Lipschitz constant $L$
(see Definition \ref{def:LP_para}). Then, $\mathtt{SlowSolveLP}$
outputs $x$ such that
\begin{align*}
c^{\top}x & \leq\min_{\ma x=b,x\geq0}c^{\top}x+\delta LR,\\
\ma x & =b,\\
x & \geq0.
\end{align*}
The algorithm takes $O(\sqrt{n}\log(nR/(\delta r)))$ Newton steps.

If we further assume that the solution $x^{*}=\arg\min_{\ma x=b,x\geq0}c^{\top}x$
is unique and that $c^{\top}x\geq c^{\top}x^{*}+\eta LR$ for any
other vertex $x$ of $\{\ma x=b,x\geq0\}$ for some $\eta>\delta\geq0$,
then we have that $\|x-x^{*}\|_{2}\leq\frac{2\delta R}{\eta}$.
\end{thm}

\begin{proof}
By Theorem \ref{thm:IPM_interior}, the point $(x_{c}^{+},x_{c}^{-},\overline{R})$
is on the central path of the modified program at $t$. After the
first call of $\mathtt{L2Step}$, we have $(\overline{x},\overline{s})$
such that $\|\overline{x}\overline{s}-t\|_{2}\leq\frac{t}{6}$ with
$t=LR$.

Theorem \ref{thm:IPM_interior} shows that $(x,s)=(x^{+}-x^{-},s^{+}-s^{\theta})\in\PriR\times\DualR$
and that $x=(1\pm\epsilon)x^{+}$ and $s=(1\pm\epsilon)s^{+}$. Since
$\epsilon=\frac{1}{100\sqrt{n}}$ and $\|x^{+}s^{+}-t\|_{2}\leq\frac{t}{6}$
with $t=LR$, we have that $\|xs-t\|_{2}\leq\frac{t}{4}$. This verifies
the condition for the second call of $\mathtt{L2Step}$.

After the second call of $\mathtt{L2Step}$, we have $(x_{\mathrm{end}},s_{\mathrm{end}})$
such that $\|x_{\mathrm{end}}s_{\mathrm{end}}-t_{\mathrm{end}}\|_{2}\leq\frac{t}{6}$
with $t_{\mathrm{end}}=\delta LR/(2n)$. Hence, by the bound on the
duality gap, 
\[
c^{\top}x_{\mathrm{end}}\leq\min_{\ma x=b,x\geq0}c^{\top}x+x_{\mathrm{end}}^{\top}s_{\mathrm{end}}\leq\min_{\ma x=b,x\geq0}c^{\top}x+2t_{\mathrm{end}}n\leq\min_{\ma x=b,x\geq0}c^{\top}x+\delta LR.
\]
For the runtime, note that $\mathtt{L2Step}$ decreases $t$ by $1-\Omega(n^{-1/2})$
factor each step. Hence, the first call takes $O(\sqrt{n}\log(nR/r))$
Newton steps and second call takes $O(\sqrt{n}\log(n/\delta))$ Newton
steps.

For the last conclusion, we assume $\delta\leq\eta$ and let $\mathcal{P}_{t}=\mathcal{P}\cap\{c^{\top}x\leq c^{\top}x^{*}+tLR\}$.
Note that $\mathcal{P}_{\eta}$ is a cone at $x^{*}$ (because there
is no vertex except $x^{*}$ with value less than $c^{\top}x^{*}+tLR$).
Hence, we have $\mathcal{P}_{\delta}-x^{*}=\frac{\delta}{\eta}(\mathcal{P}_{\eta}-x^{*})$.
Since $x\in\mathcal{P}_{\delta}$, we have that 
\[
\|x-x^{*}\|_{2}\leq\frac{\delta}{\eta}\text{diameter}(\mathcal{P}_{\eta}-x^{*})\leq\frac{2\delta R}{\eta}.
\]
\end{proof}
If we know the solution of the linear program is integral or rational
with some bound on the number of bits, then getting a solution close
enough to $x^{*}$ allows us to round the solution to an integral
solution. Therefore, the last conclusion of the theorem above gives
us an exact linear programming algorithm assuming $\ma,b,c$ are integral
and bounded. The uniqueness assumption can be achieved by perturbing
the cost vector by a random vector.

First, we show that $x^{(0)}$ defined in Theorem \ref{thm:IPM_interior}
is indeed on the central path of the modified linear program.
\begin{lem}
\label{lem:ipm_interior_modified}The modified linear program (Definition
\ref{def:IPM_interior_modified}) has an explicit central path point
$x^{(0)}=(x_{c}^{+},x_{c}^{-},\overline{R})$ at $t$.
\end{lem}

\begin{proof}
Recall that we say $(x^{+},x^{-},x^{\theta})$ is on the central path
at $t$ if $x^{+},x^{-},x^{\theta}$ are positive and it satisfies
the following equation
\begin{align}
\ma x^{+}-\ma x^{-} & =b,\nonumber \\
\sum_{i=1}^{n}x_{i}^{+}+x^{\theta} & =\widetilde{b},\nonumber \\
\ma^{\top}y+\lambda+s^{+} & =c,\label{eq:KKT_modified_LP}\\
-\ma^{\top}y+s^{-} & =\widetilde{c},\nonumber \\
\lambda+s^{\theta} & =0,\nonumber 
\end{align}
for some $s^{+},s^{-}\in\R_{>0}^{n}$, $s^{\theta}>0$, $y\in\R^{d}$
and $\lambda\in\R$.

Now, we verify the solution $x^{+}=\frac{t}{c+t/\overline{R}}$, $x^{-}=\frac{t}{c+t/\overline{R}}-x_{\circ}$,
$x^{\theta}=\overline{R}$, $x_{\circ}=\ma^{\top}(\ma\ma^{\top})^{-1}b$,
$y=0$, $s^{+}=\frac{t}{x^{+}}$, $s^{-}=\frac{t}{x^{-}}$, $s^{\theta}=\frac{t}{x^{\theta}}$,
$\lambda=-s^{\theta}$. Using $\ma x_{\circ}=b$, one can check it
satisfies all the equality constraints above. 

For the inequality constraints, using $\|c\|_{\infty}\leq L$ and
$t\geq8L\overline{R}$, we have
\begin{equation}
\frac{3}{4}\overline{R}\leq\frac{t}{L+t/\overline{R}}\leq x_{i}^{+}\leq\frac{t}{-L+t/\overline{R}}\leq\frac{3}{2}\overline{R}\label{eq:x_pos_bound}
\end{equation}
and hence $x^{+}>0$ and so is $s^{+}$. Since $\|x_{\circ}\|_{2}\leq R\leq\frac{\overline{R}}{2}$
and $x_{i}^{+}\geq\frac{3}{4}\overline{R}$ for all $i$, we have
$x_{i}^{-}\geq0$ for all $i$. Hence, $x^{-}$ and $s^{-}$ are positive.
Finally, $x^{\theta}$ and $s^{\theta}$ are positive. This proves
that $(\frac{t}{c+t/\overline{R}},\frac{t}{c+t/\overline{R}}-x_{\circ},\overline{R})$
is on the central path point at $t$.
\end{proof}
Next, we show that the near-central-path point $(x,s)$ at $t=LR$
is far from the constraints $x^{+}\geq0$ and is close to the constraints
$x^{-}\geq0$. The proof for both involves the same idea: use the
optimality condition of $x$. Throughout the rest of the section,
we are given $(x,s)\in\mathcal{P}_{\overline{R},t}\times\mathcal{D}_{\overline{R},t}$
such that $\mu=xs$ satisfies
\[
\frac{5}{6}LR\leq\mu\leq\frac{7}{6}LR.
\]
We write $\mu$ into its three parts $(\mu^{+},\mu^{-},\mu^{\theta})$.
By Lemma \ref{lem:primal_dual_representation}, we have that $x\defeq(x^{+},x^{-},x^{\theta})$
minimizes the function
\[
f(x^{+},x^{-},x^{\theta})\defeq c^{\top}x^{+}+\widetilde{c}^{\top}x^{-}-\sum_{i=1}^{n}\mu_{i}^{+}\log x_{i}^{+}-\sum_{i=1}^{n}\mu_{i}^{-}\log x_{i}^{-}-\mu^{\theta}\log x^{\theta}
\]
over the domain $\mathcal{P}_{\overline{R},t}$. The gradient of $f$
is a bit complicated. We avoid it by considering the directional derivative
at $x$ along the direction ``$v-x$'' where $v$ is the point such
that $\ma v=b$ and $v\geq r$. Since our domain is in $\mathcal{P}_{\overline{R},t}\subset\R^{2n+1}$,
we need to lift $v$ to higher dimension. So we define the point
\begin{align*}
v^{-} & =\min(x^{-},\frac{8L\overline{R}}{t}\cdot R),\\
v^{+} & =v+v^{-},\\
v^{\theta} & =\widetilde{b}-\sum_{i=1}^{n}v_{i}^{+}.
\end{align*}
First, we need to get some basic bounds on $\widetilde{b}$ and $\widetilde{c}$.
\begin{lem}
\label{lem:ipm_propert_modified}We have that $\frac{3}{4}n\overline{R}\leq\widetilde{b}\leq3n\overline{R}$
and $\widetilde{c}_{i}\geq t/(2\overline{R})$ for all $i$.
\end{lem}

\begin{proof}
By (\ref{eq:x_pos_bound}), we have $\frac{3}{4}\overline{R}\leq x_{c,i}^{+}\leq\frac{3}{2}\overline{R}$.
By the definition of $\widetilde{b}$, we have
\[
\widetilde{b}=\sum_{i=1}^{n}x_{c,i}^{+}+\overline{R}\leq\frac{3}{2}n\overline{R}+\overline{R}\leq3n\overline{R}.
\]
Similarly, we have $\widetilde{b}=\sum_{i}x_{i}^{+}+\overline{R}\geq\frac{3}{4}n\overline{R}$. 

For the bound on $\widetilde{c}$, recall that $\widetilde{c}=t/x_{c}^{-}$
with $x_{c}^{-}=x_{c}^{+}-x_{\circ}$ and $x_{\circ}=\ma^{\top}(\ma\ma^{\top})^{-1}b=\arg\min_{\ma x=b}\|x\|_{2}$.
Since we assumed the linear program has outer radius $R$, we have
that $\|x_{\circ}\|_{2}\leq R$. Hence, 
\[
x_{c,i}^{-}\leq\frac{3}{2}\overline{R}+R\leq2\overline{R}.
\]
Therefore, $\widetilde{c}\geq t/(2\overline{R})$.
\end{proof}
The following lemma shows that $(v^{+},v^{-},v^{\theta})\in\mathcal{P}_{\overline{R},t}$.
\begin{lem}
\label{lem:IPM_exact_inner_pt}We have that $(v^{+},v^{-},v^{\theta})\in\mathcal{P}_{\overline{R},t}$.
Furthermore, we have $v^{\theta}\geq\frac{1}{2}n\overline{R}$.
\end{lem}

\begin{proof}
Note that $(v^{+},v^{-},v^{\theta})$ satisfies the linear constraints
of $\mathcal{P}_{\overline{R},t}$ by construction. It suffices to
prove the vector is positive. Since $x^{-}>0$, we have $v^{-}>0$.
Since $v\geq r$, we also have $v^{+}>0$. For $v^{\theta}$, we use
$\widetilde{b}\geq\frac{3}{4}n\overline{R}$ (Lemma \ref{lem:ipm_propert_modified}),
$v\leq R$ and $v^{-}\leq\frac{8L\overline{R}}{t}\cdot R\leq R$ to
get
\begin{align*}
v^{\theta}=\widetilde{b}-\sum_{i=1}^{n}v_{i}^{+} & \geq\frac{3}{4}n\overline{R}-\sum_{i=1}^{n}(v_{i}+v_{i}^{-})\geq\frac{1}{2}n\overline{R}.
\end{align*}
\end{proof}
Next, we define the path $p(t)=(1-t)(x^{+},x^{-},x^{\theta})+t(v^{+},v^{-},v^{\theta})$.
Since $p(0)$ minimizes $f$, we have that $\frac{d}{dt}f(p(t))|_{t=0}\geq0$.
In particular, we have
\begin{align}
0 & \leq\frac{d}{dt}f(p(t))|_{t=0}\nonumber \\
 & =c^{\top}(v^{+}-x^{+})+\widetilde{c}^{\top}(v^{-}-x^{-})-\sum_{i=1}^{n}\frac{\mu_{i}^{+}}{x_{i}^{+}}(v^{+}-x^{+})_{i}-\sum_{i=1}^{n}\frac{\mu_{i}^{-}}{x_{i}^{-}}(v^{-}-x^{-})_{i}-\frac{\mu^{\theta}}{x^{\theta}}(v_{\theta}-x^{\theta})\nonumber \\
 & =\frac{\mu^{\theta}}{x^{\theta}}(x^{\theta}-v_{\theta})+\sum_{i=1}^{n}(c_{i}-\frac{\mu_{i}^{+}}{x_{i}^{+}})(v^{+}-x^{+})_{i}+\sum_{i=1}^{n}(\widetilde{c}_{i}-\frac{\mu_{i}^{-}}{x_{i}^{-}})(v^{-}-x^{-})_{i}.\label{eq:LP_exact_3_term}
\end{align}
Now, we bound each term one by one. For the first term, we note that
\begin{equation}
\frac{\mu^{\theta}}{x^{\theta}}(x^{\theta}-v_{\theta})\leq\mu^{\theta}\leq2LR.\label{eq:LP_exact_3_term_1}
\end{equation}
For the second term in (\ref{eq:LP_exact_3_term}), we have the following
\begin{lem}
\label{lem:LP_exact_3_term_1}We have that $\sum_{i=1}^{n}(c_{i}-\frac{\mu_{i}^{+}}{x_{i}^{+}})(v^{+}-x^{+})_{i}\leq4nL\overline{R}-\frac{LRr}{2\min_{i}x_{i}^{+}}$.
\end{lem}

\begin{proof}
Note that
\begin{align*}
\sum_{i=1}^{n}(c_{i}-\frac{\mu_{i}^{+}}{x_{i}^{+}})(v^{+}-x^{+})_{i} & =\sum_{i=1}^{n}(c_{i}v_{i}^{+}-\frac{\mu_{i}^{+}}{x_{i}^{+}}v_{i}^{+}-c_{i}x_{i}^{+}+\mu_{i}^{+})\\
 & \leq\sum_{i=1}^{n}c_{i}v_{i}^{+}+\sum_{i=1}^{n}\mu_{i}^{+}-\sum_{i=1}^{n}\frac{\mu_{i}^{+}}{x_{i}^{+}}v_{i}^{+}\\
 & \leq\|c\|_{\infty}\|v^{+}\|_{1}+2nLR-\frac{1}{2}\sum_{i=1}^{n}\frac{LRr}{x_{i}^{+}}
\end{align*}
where we used $\mu_{i}^{+}\in[\frac{LR}{2},2LR]$ and $v_{i}^{+}\geq v_{i}\geq r$
at the end. The result follows from $\|c\|_{\infty}\leq L$, $\|v^{+}\|_{1}\leq\widetilde{b}\leq3n\overline{R}$
(Lemma \ref{lem:ipm_propert_modified}).
\end{proof}
For the third term in (\ref{eq:LP_exact_3_term}), we have the following
\begin{lem}
\label{lem:LP_exact_3_term_2}We have that $\sum_{i=1}^{n}(\widetilde{c}_{i}-\frac{\mu_{i}^{-}}{x_{i}^{-}})(v^{-}-x^{-})_{i}\leq2LR-\frac{t}{4\overline{R}}\max_{i}x_{i}^{-}$.
\end{lem}

\begin{proof}
Using $v^{-}=\min(x^{-},\frac{8L\overline{R}}{t}\cdot R)$, we have
$v_{i}^{-}\leq x_{i}^{-}$. We can ignore the terms with $v_{i}^{-}=x_{i}^{-}$. 

For $v_{i}^{-}<x_{i}^{-}$, we have $x_{i}^{-}\geq\frac{8L\overline{R}}{t}R$.
Using $\widetilde{c}_{i}\geq\frac{t}{2\overline{R}}$ (Lemma \ref{lem:ipm_propert_modified}),
we have
\[
\widetilde{c}_{i}-\frac{\mu_{i}^{-}}{x_{i}^{-}}\geq\widetilde{c}_{i}-\frac{\mu_{i}^{-}}{\frac{8L\overline{R}}{t}R}\geq\widetilde{c}_{i}-\frac{2LR}{\frac{8L\overline{R}}{t}R}=\widetilde{c}_{i}-\frac{t}{4\overline{R}}\geq\frac{t}{4\overline{R}}.
\]
Hence, we have
\[
\sum_{i=1}^{n}(\widetilde{c}_{i}-\frac{\mu_{i}^{-}}{x_{i}^{-}})(v^{-}-x^{-})_{i}\leq\frac{t}{4\overline{R}}\sum_{i=1}^{n}(v^{-}-x^{-})_{i}\leq\frac{t}{4\overline{R}}(\frac{8L\overline{R}}{t}\cdot R-\max_{i}x_{i}^{-}).
\]
\end{proof}
Combining (\ref{eq:LP_exact_3_term}), (\ref{eq:LP_exact_3_term_1}),
Lemma \ref{lem:LP_exact_3_term_1} and Lemma \ref{lem:LP_exact_3_term_2},
we have
\begin{align*}
0 & \leq\frac{\mu^{\theta}}{x^{\theta}}(x^{\theta}-v_{\theta})+\sum_{i=1}^{n}(c_{i}-\frac{\mu_{i}^{+}}{x_{i}^{+}})(v^{+}-x^{+})_{i}+\sum_{i=1}^{n}(\widetilde{c}_{i}-\frac{\mu_{i}^{-}}{x_{i}^{-}})(v^{-}-x^{-})_{i}\\
 & \leq2LR+4nL\overline{R}-\frac{LRr}{2\min_{i}x_{i}^{+}}+2LR-\frac{t}{4\overline{R}}\max_{i}x_{i}^{-}\\
 & =5nL\overline{R}-\frac{LRr}{2\min_{i}x_{i}^{+}}-\frac{t}{4\overline{R}}\max_{i}x_{i}^{-}.
\end{align*}
Hence, we show that $(x^{+},x^{-},x^{\theta})$ is close to the central
path at $t=LR$ implies that $\min_{i}x_{i}^{+}$ cannot be too small
and $\max_{i}x_{i}^{-}$ cannot be too large
\[
\frac{LRr}{2\min_{i}x_{i}^{+}}+\frac{t}{4\overline{R}}\max_{i}x_{i}^{-}\leq5nL\overline{R}.
\]
In particular, this shows the following:
\begin{lem}
\label{lem:LP_exact_distance_OPT}We have that $\min_{i}x_{i}^{+}\geq\frac{Rr}{10n\overline{R}}$
and $\max_{i}x_{i}^{-}\leq\frac{20nL\overline{R}}{t}\cdot\overline{R}$.
\end{lem}

Now, we are ready to prove the second conclusion of Theorem \ref{thm:IPM_interior}.
\begin{lem}
\label{lemma:IPM_center_original}For any primal $x\defeq(x^{+},x^{-},x^{\theta})\in\mathcal{P}_{\overline{R},t}$
and dual $s\defeq(s^{+},s^{-},s^{\theta})\in\mathcal{D}_{\overline{R},t}$
such that $\frac{5}{6}LR\leq x_{i}s_{i}\leq\frac{7}{6}LR$, we have
that
\[
(x^{+}-x^{-},s^{+}-s^{\theta})\in\PriR\times\DualR
\]
and that $x_{i}^{-}\leq\epsilon x_{i}^{+}$ and $s^{\theta}\leq\epsilon s_{i}^{+}$
for all $i$.
\end{lem}

\begin{proof}
First we check $x\defeq x^{+}-x^{-}\in\PriR$. By the choice of $\overline{R}$
and $t$, Lemma \ref{lem:LP_exact_distance_OPT} shows that 
\[
\frac{\max_{i}x_{i}^{-}}{\min_{i}x_{i}^{+}}\leq\frac{\frac{20nL\overline{R}}{t}\cdot\overline{R}}{\frac{Rr}{10n\overline{R}}}=\frac{200n^{2}L\frac{\overline{R}^{3}}{Rr}}{t}\leq\epsilon.
\]
Hence, we have $x^{+}-x^{-}>0$ and that $\ma(x^{+}-x^{-})=b$. 

Next, we check $s\defeq s^{+}-s^{\theta}\in\DualR$. Since $x\in\mathcal{P}$,
we have $x\leq R$ and $x_{i}^{+}\leq\frac{3}{2}x_{i}\leq\frac{3}{2}R$.
Since $x_{i}^{+}s_{i}^{+}\geq\frac{5}{6}LR$, we have $s_{i}^{+}\geq\frac{1}{2}L.$
On the other hand, we have $x^{\theta}=\widetilde{b}-\sum_{i=1}^{n}x_{i}^{+}\geq\widetilde{b}-2nR\geq\frac{1}{2}n\overline{R}$
(Lemma \ref{lem:ipm_propert_modified}). Hence, $s^{\theta}\leq\frac{\frac{7}{6}LR}{\frac{1}{2}n\overline{R}}\leq\frac{5LR}{2n\overline{R}}.$
Combining both and the choice of $\overline{R}$, we have
\[
\frac{s^{\theta}}{\min_{i}s_{i}^{+}}\leq\frac{\frac{5LR}{2n\overline{R}}}{L/2}=\frac{5R}{n\overline{R}}\leq\epsilon
\]
Hence, we have $s=s^{+}-s^{\theta}>0$ and that $\ma^{\top}y+s=\ma^{\top}y+s^{+}-s^{\theta}=\ma^{\top}y+s^{+}+\lambda=c$
(See (\ref{eq:KKT_modified_LP})).
\end{proof}

\subsection{Why $\sqrt{n}$?}

The central path is the solution to the following ODE
\begin{align*}
S_{t}\frac{d}{dt}x_{t}+X_{t}\frac{d}{dt}s_{t} & =1,\\
A\frac{d}{dt}x_{t} & =0,\\
A^{\top}\frac{d}{dt}y_{t}+\frac{d}{dt}s_{t} & =0.
\end{align*}
Solving this linear system, we have that $S_{t}\frac{dx_{t}}{dt}=(I-P_{t})1$
and $X_{t}\frac{ds_{t}}{dt}=P_{t}1$ where $P_{t}=X_{t}A^{\top}(AS_{t}^{-1}X_{t}A^{\top})^{-1}AS_{t}^{-1}$.
Using that $x_{t}s_{t}=t$, we have that
\[
P_{t}=X_{t}A^{\top}(AX_{t}^{2}A^{\top})^{-1}AX_{t}=S_{t}^{-1}A^{\top}(AS_{t}^{-2}A^{\top})^{-1}AS_{t}^{-1}
\]
and that $X_{t}^{-1}\frac{dx_{t}}{dt}=\frac{1}{t}(I-P_{t})1$ and
$S_{t}^{-1}\frac{ds_{t}}{dt}=\frac{1}{t}P_{t}1$. Equivalently, we
have
\[
\frac{d\ln x_{t}}{d\ln t}=(I-P_{t})1\text{ and }\frac{d\ln s_{t}}{d\ln t}=P_{t}1.
\]
Note that 
\[
\norm{P_{t}1}_{\infty}\leq\norm{P_{t}1}_{2}=\sqrt{n}.
\]
Hence, $x_{t}$ and $s_{t}$ can change by at most a constant factor
when we change $t$ by a $1\pm\frac{1}{\sqrt{n}}$ factor.
\begin{xca}
If we are given $x$ such that $\norm{\ln x-\ln x_{t}}_{\infty}=O(1)$,
then we can find $x_{t}$ by solving $\tilde{O}(1)$ linear systems.

\end{xca}


\section{Interior Point Method for Convex Programs}

The interior point method can be used to optimize any convex function.
For more in-depth treatment, please see the structural programming
section in \cite{nesterov1998introductory}.

Recall that any convex optimization problem
\[
\min_{x}f(x)
\]
can be rewritten in the epigraph form as
\[
\min_{\{(x,t):\ f(x)\leq t\}}t.
\]
Hence, it suffices to study the problem $\min_{x\in K}c^{\top}x$.
Similar to the case of linear programs, we replace the hard constraint
$x\in K$ by a soft constraint as follows:
\[
\min_{x}\phi_{t}(x)\text{ where }\phi_{t}(x)=tc^{\top}x+\phi(x).
\]
where $\phi(x)$ is a convex function such that $\phi(x)\rightarrow+\infty$
as $x\rightarrow\partial K$. Note that we put the parameter $t$
in front of the cost $c^{\top}x$ instead of $\phi$ as in the last
lecture, it is slightly more convenient here. We say $\phi$ is a
\emph{barrier} for $K$. To be concrete, we can always keep in mind
$\phi(x)=-\sum_{i=1}^{n}\ln x_{i}$. As before, we define the central
path.
\begin{defn}
The central path $x_{t}=\arg\min_{x}\phi_{t}(x)$.
\end{defn}

The interior point method follows the following framework:
\begin{enumerate}
\item Find $x$ close to $x_{0}$.
\item While $t$ is not large enough,
\begin{enumerate}
\item Move $x$ closer to $x_{t}$
\item $t\rightarrow(1+h)\cdot t$.
\end{enumerate}
\end{enumerate}

\subsection{Self-concordance}

In this section, we give a general analysis for the Newton method.
In the next section, we will use this to show that interior point
method can be generalized to convex optimization. A key property of
the Newton method is that it is invariant under linear transformation.
In general, whenever a method uses $k^{th}$ order information, we
need to assume the $k^{th}$ derivative is continuous. Otherwise,
the $k^{th}$ derivative is not useful for algorithmic purposes. For
the Newton method, it is convenient to assume that the Hessian is
Lipschitz. Since the method is invariant under linear transformation,
it only makes sense to impose an assumption that is invariant under
linear transformation.
\begin{defn}
Given a convex function $f:\R^{n}\rightarrow\R$, and any point $x\in\R^{n},$
define the norm $\norm ._{x}$ as 
\[
\norm v_{x}^{2}=v^{\top}\nabla^{2}f(x)v.
\]
We call a function $f$ self-concordant if for any $h\in\Rn$ and
any $x$ in $\dom f$, we have
\[
|D^{3}f(x)[h,h,h]|\leq2\norm h_{x}^{3}
\]
where $D^{k}f(x)[h_{1},h_{2},\cdots,h_{k}]$ is the directional $k^{th}$
derivative of $f$ along the directions $h_{1},h_{2},\cdots,h_{k}$.
\end{defn}

\begin{rem*}
The constant $2$ is chosen so that $-\ln(x)$ exactly satisfies the
assumption and it is not very important, in that by scaling $f$,
we can change any constant to any other constant.
\end{rem*}
\begin{xca}
Show that the following property is equivalent to self-concordance
as defined above: restricted on any straight line $g(t)=f(x+th)$,
we have $g'''(t)\leq2g''(t)^{3/2}$.
\end{xca}

\begin{xca}
Show that the functions $x^{\top}Ax$, $-\ln x$, $-\ln(1-\sum x_{i}^{2})$,
$-\ln\det X$ are self-concordant under suitable nonnegativity conditions.
\end{xca}

The self-concordance condition says that locally, the Hessian does
not change too fast, i.e., the change in the Hessian is bounded by
its magnitude (to the power $1.5$). We will skip the proof of the
lemma below.
\begin{lem}
Given a self-concordant function $f$, for any $h_{1},h_{2},h_{3}\in\Rn$,
we have 
\[
D^{3}f(x)[h_{1},h_{2},h_{3}]\leq2\norm{h_{1}}_{x}\norm{h_{2}}_{x}\norm{h_{3}}_{x}.
\]
\end{lem}

From the self-concordance condition, we have the following more directly
usable property.
\begin{lem}
\label{lem:self_con_global}For a self-concordant function $f$ and
any $x\in\dom f$ and any $\norm{y-x}_{x}<1$, we have that
\[
(1-\norm{y-x}_{x})^{2}\nabla^{2}f(x)\preceq\nabla^{2}f(y)\preceq\frac{1}{(1-\norm{y-x}_{x})^{2}}\nabla^{2}f(x).
\]
\end{lem}

\begin{proof}
Let $\alpha(t)=\left\langle \nabla^{2}f(x+t(y-x))u,u\right\rangle $.
Then, we have that 
\[
\alpha'(t)=D^{3}f(x+t(y-x))[y-x,u,u].
\]
By self-concordance, we have
\begin{equation}
\left|\alpha'(t)\right|\leq2\norm{y-x}_{x+t(y-x)}\norm u_{x+t(y-x)}^{2}.\label{eq:self_con_global_alpha}
\end{equation}

For $u=y-x$, we have $\left|\alpha'(t)\right|\leq2\alpha(t)^{\frac{3}{2}}$.
Hence, we have $\frac{d}{dt}\frac{1}{\sqrt{\alpha(t)}}\geq-1.$ Integrating
both sides w.r.t. $t$, we have
\[
\frac{1}{\sqrt{\alpha(t)}}\geq\frac{1}{\sqrt{\alpha(0)}}-t=\frac{1}{\norm{x-y}_{x}}-t.
\]
Rearranging it gives
\[
\norm{y-x}_{x+t(y-x)}^{2}\leq\frac{1}{(\frac{1}{\norm{x-y}_{x}}-t)^{2}}=\frac{\norm{x-y}_{x}^{2}}{(1-t\norm{x-y}_{x})^{2}}.
\]

Going back to general $u$, (\ref{eq:self_con_global_alpha}) gives
\[
\left|\alpha'(t)\right|\leq2\frac{\norm{x-y}_{x}}{1-t\norm{x-y}_{x}}\alpha(t).
\]
Rearranging,
\[
\left|\frac{d}{dt}\ln\alpha(t)\right|\leq2\frac{\norm{x-y}_{x}}{1-t\norm{x-y}_{x}}=-2\frac{d}{dt}\ln(1-t\norm{x-y}_{x})
\]
Integrating both from $t=0$ to $1$ gives the result.
\end{proof}

\subsection{Self-concordant barrier functions}

To analyze the algorithm above, we need to assume that $\phi$ is
well-behaved. We measure the quality of $\phi$ by $\|\nabla\phi\|_{\nabla^{2}\phi(x)^{-1}}$.
One can think this as the Lipschitz constant of $\phi$ but measured
in the local norm. 
\begin{defn}
We call $\phi$ a $\nu$-self-concordant barrier for $K$ if $\phi$
is self-concordant, $\phi(x)\rightarrow+\infty$ as $x\rightarrow\partial K$
and that $\|\nabla\phi(x)\|_{\nabla^{2}\phi(x)^{-1}}^{2}\leq\nu$
for all $x$.
\end{defn}

Not all convex functions are self-concordant. However, for our purpose,
it suffices to show that we can construct a self-concordant barrier
for any convex set.
\begin{thm}
Any convex set has an $n$-self concordant barrier.
\end{thm}

Unfortunately, this is an existence result and the barrier function
is expensive to compute. In practice, we construct self-concordant
barriers out of simpler ones:
\begin{lem}
The following functions are self-concordant barriers. We use $\nu$-sc
as a short form for $\nu$-self-concordant barrier.
\begin{itemize}
\item $-\ln x$ is $1$-sc for $\{x\geq0\}$.
\item $-\ln\cos(x)$ is $1$-sc for $\{\left|x\right|\leq\frac{\pi}{2}\}$.
\item $-\ln(t^{2}-\norm x^{2})$ is $2$-sc for $\{t\geq\norm x_{2}\}$.
\item $-\ln\det X$ is $n$-sc for $\{X\in\R^{n\times n},X\succeq0\}$.
\item $-\ln x-\ln(\ln x+t)$ is $2$-sc for $\{x\geq0,t\geq-\ln x\}$.
\item $-\ln t-\ln(\ln t-x)$ is $2$-sc for $\{t\geq e^{x}\}$.
\item $-\ln x-\ln(t-x\ln x)$ is $2$-sc for $\{x\geq0,t\geq x\ln x\}$.
\item $-2\ln t-\ln(t^{2/p}-x^{2})$ is $4$-sc for $\{t\geq|x|^{p}\}$ for
$p\geq1$.
\item $-\ln x-\ln(t^{p}-x)$ is $2$-sc for $\{t^{p}\geq x\geq0\}$ for
$0<p\leq1$.
\item $-\ln t-\ln(x-t^{-1/p})$ is $2$-sc for $\{x>0,t\geq x^{-p}\}$ for
$p\geq1$.
\item $-\ln x-\ln(t-x^{-p})$ is $2$-sc for $\{x>0,t\geq x^{-p}\}$ for
$p\geq1$.
\end{itemize}
\end{lem}

The following lemma shows how we can combine barriers.
\begin{lem}
If $\phi_{1}$ and $\phi_{2}$ are $\nu_{1}$ and $\nu_{2}$-self
concordant barriers for $K_{1}$ and $K_{2}$ respectively, then $\phi_{1}+\phi_{2}$
is a $\nu_{1}+\nu_{2}$ self concordant barrier for $K_{1}\cap K_{2}$.
\end{lem}

\begin{lem}
If $\phi$ is a $\nu$-self concordant barrier for $K$, then $\phi(Ax+b)$
is $\nu$-self concordant for $\{y:Ay+b\in K\}$.
\end{lem}

\begin{xca}
Using the lemmas above, prove that $-\sum_{i=1}^{m}\ln(a_{i}^{\top}x-b_{i})$
is an $m$-self concordant barrier for the convex set $\{Ax\geq b\}$. 
\end{xca}


\subsection{Convergence of Newton method for self-concordant functions}

Now we are ready to study the convergence of Newton method.
\begin{lem}
\label{lem:Newton_improvement}Given a self-concordant convex function
$f$, consider the iteration 
\[
x'=x-(\nabla^{2}f(x))^{-1}\nabla f(x).
\]
Suppose that $r=\norm{\nabla f(x)}_{\nabla^{2}f(x)^{-1}}<1$, then
we have
\[
\norm{\nabla f(x')}_{\nabla^{2}f(x')^{-1}}\leq\frac{r^{2}}{(1-r)^{2}}.
\]
\end{lem}

\begin{rem}
Note that $\norm{\nabla f(x)}_{\nabla^{2}f(x)^{-1}}=\norm{\nabla^{2}f(x)^{-1}\nabla f(x)}_{x}$
is the step size of the Newton method. This is a measurement of the
error, since the goal is to find $x$ with $\nabla f(x)=0$.
\end{rem}

\begin{proof}
Lemma \ref{lem:self_con_global} shows that
\[
\nabla^{2}f(x')\succeq(1-r)^{2}\nabla^{2}f(x).
\]
and hence
\[
\norm{\nabla f(x')}_{\nabla^{2}f(x')^{-1}}\leq\frac{\norm{\nabla f(x')}_{\nabla^{2}f(x)^{-1}}}{(1-r)}.
\]
To bound $\nabla f(x')$, we calculate that
\begin{align}
\nabla f(x') & =\nabla f(x)+\int_{0}^{1}\nabla^{2}f(x+t(x'-x))(x'-x)dt\nonumber \\
 & =\nabla f(x)-\int_{0}^{1}\nabla^{2}f(x+t(x'-x))(\nabla^{2}f(x))^{-1}\nabla f(x)dt\nonumber \\
 & =\left(\nabla^{2}f(x)-\int_{0}^{1}\nabla^{2}f(x+t(x'-x))dt\right)(\nabla^{2}f(x))^{-1}\nabla f(x).\label{eq:grad_f_sc}
\end{align}
For the term in the bracket, we use Lemma \ref{lem:self_con_global}
to get (note that $\int_{0}^{1}(1-tr)^{2}dt=1-r+\frac{1}{3}r^{2}$):
\[
(1-r+\frac{1}{3}r^{2})\nabla^{2}f(x)\preceq\int_{0}^{1}\nabla^{2}f(x+t(x'-x))dt\preceq\frac{1}{1-r}\nabla^{2}f(x).
\]
Therefore, we have
\[
\norm{(\nabla^{2}f(x))^{-\frac{1}{2}}\left(\nabla^{2}f(x)-\int_{0}^{1}\nabla^{2}f(x+t(x'-x))dt\right)(\nabla^{2}f(x))^{-\frac{1}{2}}}_{\text{op}}\leq\max\left(\frac{r}{1-r},r-\frac{1}{3}r^{2}\right)=\frac{r}{1-r}.
\]
Putting it into (\ref{eq:grad_f_sc}) gives
\begin{align*}
\norm{\nabla f(x')}_{\nabla^{2}f(x)^{-1}} & =\norm{\nabla^{2}f(x)^{-\frac{1}{2}}\nabla f(x')}_{2}\\
 & \le\norm{(\nabla^{2}f(x))^{-\frac{1}{2}}\left(\nabla^{2}f(x)-\int_{0}^{1}\nabla^{2}f(x+t(x'-x))dt\right)(\nabla^{2}f(x))^{-\frac{1}{2}}}_{\text{op}}\norm{(\nabla^{2}f(x))^{-\frac{1}{2}}\nabla f(x)}_{2}\\
 & \leq\frac{r}{1-r}\cdot r\\
 & =\frac{r^{2}}{1-r}.
\end{align*}
\end{proof}
Finally, we bound the error of the current iterate in terms of $\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}$.
\begin{lem}
\label{lem:self_concordant_err}Given $x$ such that $\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}\leq\frac{1}{6}$,
we have that
\begin{itemize}
\item $\|x-x^{*}\|_{x^{*}}\leq2\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}$,
\item $\|x-x^{*}\|_{x}\leq\frac{4}{3}\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}$,
\item $f(x)\leq f(x^{*})+\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}^{2}.$
\end{itemize}
\end{lem}

\begin{proof}
Let $r=\|x-x^{*}\|_{x}$. Suppose that $r\leq\frac{1}{4}$. Note that
\[
\nabla f(x)=\nabla f(x)-\nabla f(x^{*})=\int_{0}^{1}\nabla^{2}f(x^{*}+t(x-x^{*}))(x-x^{*})dt.
\]
Using that $\nabla^{2}f(x^{*}+t(x-x^{*}))\succeq(1-(1-t)r)^{2}\nabla^{2}f(x)$
(Lemma \ref{lem:self_con_global}), we have
\begin{align*}
\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}} & =\norm{\int_{0}^{1}\nabla^{2}f(x^{*}+t(x-x^{*}))(x-x^{*})\,dt}_{\nabla^{2}f(x)^{-1}}\\
 & \geq\int_{0}^{1}(1-(1-t)r)^{2}\|x-x^{*}\|_{x}dt\\
 & =\left(1-r+\frac{r^{2}}{3}\right)r\geq\frac{3r}{4}.
\end{align*}

Using $\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}\leq\frac{1}{6}$, we
have indeed $r\leq\frac{1}{4}$ (our lower bound above is a non-decreasing
function). Using Lemma \ref{lem:self_con_global} again, we have
\[
\|x-x^{*}\|_{x^{*}}\leq\frac{\|x-x^{*}\|_{x}}{1-r}\leq\frac{4}{3}\frac{1}{1-\frac{1}{4}}\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}\leq2\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}.
\]

For the bound for $f(x)$, we have that
\[
f(x)=f(x^{*})+\left\langle \nabla f(x^{*}),x-x^{*}\right\rangle +\int_{0}^{1}(1-t)(x-x^{*})^{\top}\nabla^{2}f(x^{*}+t(x-x^{*}))(x-x^{*})dt.
\]
Using that $\nabla^{2}f(x^{*}+t(x-x^{*}))\preceq\frac{1}{(1-(1-t)r)^{2}}\nabla^{2}f(x)$,
we have
\begin{align*}
f(x) & \leq f(x^{*})+\int_{0}^{1}\frac{1-t}{(1-(1-t)r)^{2}}dt\cdot\|x-x^{*}\|_{x}^{2}\\
 & =f(x^{*})+\frac{1}{r^{2}}(\frac{r}{1-r}+\log(1-r))\|x-x^{*}\|_{x}^{2}\\
 & \leq f(x^{*})+(\frac{1}{2}+r)\|x-x^{*}\|_{x}^{2}\\
 & \le f(x^{*})+\norm{\nabla f(x)}_{\nabla^{2}f(x)^{-1}}^{2}
\end{align*}
where we used $r\leq\frac{1}{4}$ at the end.
\end{proof}

\subsection{Main Algorithm and Analysis}

\begin{algorithm2e}[H]

\caption{$\mathtt{InteriorPointMethod}$}

\SetAlgoLined

\textbf{Input:} A $\nu$-self-concordant barrier $\phi$ for $K$,
the minimizer $x$ of $\phi$. 

Define $f_{t}(x)=tc^{\top}x+\phi(x)$. $t=\frac{1}{6}\|c\|_{\nabla^{2}\phi(x)^{-1}}^{-1}$.

\While{$t\leq\frac{\nu+\sqrt{\nu}}{\epsilon}$}{

$x\leftarrow x-\nabla^{2}f_{t}(x)^{-1}\nabla f_{t}(x)$.

$t\leftarrow(1+h)t$ with $h=\frac{1}{9\sqrt{\nu}}$.

}

\Return $x$.

\end{algorithm2e}

We first explain the initialization and termination conditions. We
start with the point $x^{(0)}=\arg\min\phi(x)$ (or a point very close
to it). The choice of $t_{0}$ is to ensure that the norm of the initial
gradient in the inverse Hessian norm is bounded by a convenient constant
less than $1$ (we use $1/6$). 

For the termination, intuitively, we can see that $\min f_{t}(x_{t})$
tends to optimality as $t\rightarrow\infty$. We first need a lemma
showing that the gradient of $\phi$ is small.
\begin{lem}[Duality Gap]
Suppose that $\phi$ is a $\nu$-self concordant barrier. For any
$x,y\in K$, we have that
\[
\left\langle \nabla\phi(x),y-x\right\rangle \leq\nu.
\]
\end{lem}

\begin{proof}
Let $\alpha(t)=\left\langle \nabla\phi(z_{t}),y-x\right\rangle $
where $z_{t}=x+t(y-x)$. Then, we have
\[
\alpha'(t)=\left\langle \nabla^{2}\phi(z_{t})(y-x),y-x\right\rangle .
\]
Note that
\[
\alpha(t)\leq\norm{\nabla\phi(z_{t})}_{\nabla^{2}\phi(z_{t})^{-1}}\norm{y-x}_{\nabla^{2}\phi(z_{t})}\leq\sqrt{\nu}\norm{y-x}_{\nabla^{2}\phi(z_{t})}.
\]
Hence, we have $\alpha'(t)\geq\frac{1}{\nu}\alpha(t)^{2}.$ If $\alpha(0)\leq0$,
then we are done. Otherwise, $\alpha$ is increasing and hence, $\alpha(1)>0$.
Therefore, 
\[
\frac{1}{\alpha(1)}\leq\frac{1}{\alpha(0)}-\frac{1}{v}
\]
and we have $\alpha(0)\leq\nu$.
\end{proof}
\begin{lem}[Duality Gap]
 Suppose that $\phi$ is a $\nu$-self concordant barrier, we have
that
\[
\left\langle c,x_{t}\right\rangle \leq\left\langle c,x^{*}\right\rangle +\frac{\nu}{t},
\]
where $x_{t}=\argmin f_{t}(x)$.\\
More generally, for any $x$ such that $\|tc+\nabla\phi(x)\|_{(\nabla^{2}\phi(x))^{-1}}\leq\frac{1}{6}$,
we have that
\[
\left\langle c,x\right\rangle \leq\left\langle c,x^{*}\right\rangle +\frac{\nu+\sqrt{\nu}}{t}.
\]
\end{lem}

\begin{proof}
Let $x^{*}$ be a minimizer of $c^{\top}x$ on $K$. By optimality
of $x_{t}$for $f_{t}$, we have $tc+\nabla\phi(x_{t})=0$. Therefore,
we have 
\[
\left\langle c,x_{t}\right\rangle -\left\langle c,x^{*}\right\rangle =\frac{1}{t}\left\langle \nabla\phi(x_{t}),x^{*}-x_{t}\right\rangle \leq\frac{\nu}{t}.
\]

For the second result, let $f(x)=tc^{T}x+\phi(x)$. Lemma \ref{lem:self_concordant_err}
shows that
\[
\|x-x_{t}\|_{x}\leq2\|\nabla f(x)\|_{\nabla^{2}f(x)^{-1}}\leq\frac{1}{3}.
\]
Hence,
\begin{align*}
\left\langle c,x-x_{t}\right\rangle  & \leq\|c\|_{\nabla^{2}\phi(x)^{-1}}\|x-x_{t}\|_{x}\leq\frac{1}{3}\|c\|_{\nabla^{2}\phi(x)^{-1}}
\end{align*}
Using $c=\frac{tc+\nabla\phi(x)}{t}-\frac{\nabla\phi(x)}{t}$, we
have
\begin{align*}
\left\langle c,x-x_{t}\right\rangle  & \leq\frac{1}{3t}\left(\|tc+\nabla\phi(x)\|_{\nabla^{2}\phi(x)^{-1}}+\|\nabla\phi(x)\|_{\nabla^{2}\phi(x)^{-1}}\right)\\
 & \leq\frac{1}{3t}\left(\frac{1}{6}+\sqrt{\nu}\right)\leq\frac{\sqrt{\nu}}{t}.
\end{align*}
This gives the result.
\end{proof}
Hence, it suffices to end with $t=(\nu+\sqrt{\nu})/\varepsilon$,
which is exactly same as the previous lecture.
\begin{thm}
Given a $\nu$-self concordant barrier $\phi$ and its minimizer.
We can find $x\in K$ such that $c^{\top}x\leq c^{\top}x^{*}+\epsilon$
in 
\[
O(\sqrt{\nu}\log(\frac{\nu}{\epsilon}\|c\|_{\nabla^{2}\phi(x^{(0)})^{-1}}))
\]
iterations.
\end{thm}

\begin{proof}
We prove by induction that $\|\nabla f_{t}(x)\|_{\nabla^{2}f_{t}(x))^{-1}}\leq\frac{1}{6}$
at the beginning of each iteration. This is true at the beginning
by the definition of initial $t$. By Lemma \ref{lem:Newton_improvement},
after the Newton step, we have
\[
\|\nabla f_{t}(x)\|_{\nabla^{2}f_{t}(x)^{-1}}\leq\left(\frac{1/6}{1-1/6}\right){}^{2}=\frac{1}{25}.
\]
Let $t'=(1+h)t$ with $h=\frac{1}{9\sqrt{\nu}}$. Note that $\nabla f_{t'}(x)=(1+h)tc+\nabla\phi(x)$
and hence $\nabla f_{t'}(x)=(1+h)\nabla f_{t}(x)-h\nabla\phi(x)$
and $\nabla^{2}f_{t'}(x)=\nabla^{2}f_{t}(x)=\nabla^{2}\phi(x).$ Therefore,
we have that
\begin{align*}
\|\nabla f_{t'}(x)\|_{\nabla^{2}f_{t'}(x)^{-1}} & =\|(1+h)\nabla f_{t}(x)-h\nabla\phi(x)\|_{\nabla^{2}f_{t}(x)^{-1}}\\
 & \leq(1+h)\|\nabla f_{t}(x)\|_{\nabla^{2}f_{t}(x)^{-1}}+h\|\nabla\phi(x)\|_{\nabla^{2}\phi(x)^{-1}}\\
 & \leq\frac{1+h}{25}+h\sqrt{\nu}\leq\frac{1}{6}.
\end{align*}

This completes the induction.
\end{proof}

