
\section{Hamiltonian Monte Carlo}

Hamiltonian dynamics is an alternative way to formulate Newtonian
mechanics. The Hamiltonian $H$ captures both the potential and kinetic
energy of a particle as a function of its position and velocity. The
dynamics can be described by the following differential equations.
\begin{align*}
\frac{dx}{dt} & =\frac{\partial H(x,v)}{\partial v},\\
\frac{dv}{dt} & =-\frac{\partial H(x,v)}{\partial x}.
\end{align*}
These equations preserve the Hamiltonian function $H$. In the simplest
Euclidean setting, it can be defined as follows.
\[
H(x,v)=f(x)+\frac{1}{2}\norm v^{2}
\]
so that 
\[
\frac{dx}{dt}=v,\quad\frac{dv}{dt}=-\nabla f(x)
\]
or
\[
\frac{d^{2}x}{dt^{2}}=-\nabla f(x).
\]

More generally, the Hamiltonian can depend on a function that defines
a local metric:
\[
H(x,v)=f(x)+\frac{1}{2}\log((2\pi)^{n}\det g(x))+\frac{1}{2}v^{T}g(x)^{-1}v
\]
where $g(x)$ is a matrix, and when it is PSD, it defines a local
norm at $x$. In this sense, we can view the dynamics as evolving
on a manifold with local metric $g(x)$. Here in this chapter, we
will focus on the case when $g(x)=I$, the standard Euclidean metric.

(Riemannian) Hamiltonian Monte Carlo (or RHMC) \cite{Neal1996,Neal2012}\cite{Girolami2009,Girolami2011}
is a Markov Chain Monte Carlo method for sampling from a desired distribution.
Each step of the method consists of the following: At a current point
$x$,
\begin{enumerate}
\item Pick a random velocity $y$ according to a local distribution defined
by $x$ (in the simplest setting, this is the standard Gaussian distribution
for every $x$).
\item Move along the Hamiltonian curve defined by Hamiltonian dynamics at
$(x,y)$ for time (distance) $\delta$.
\end{enumerate}
For the choice of $H$ above, the marginal distribution of the current
point $x$ approaches the target distribution with density proportional
to $e^{-f}$. Note that HMC does not require a Metropolis filter!
Thus, unlike the walks we have seen so far, its step sizes are not
limited by this consideration even in high dimension. Hamiltonian
Monte Carlo can be used for sampling from a general distribution $e^{-H(x,y)}$. 
\begin{defn}
Given a continuous, twice-differentiable function $H:\mathcal{M}\times\Rn\subset\R^{n}\times\R^{n}\rightarrow\R$
(\emph{Hamiltonian}) where $\mathcal{M}$ is the $x$ domain of $H$,
we say $(x(t),y(t))$ follows a \emph{Hamiltonian curve} if it satisfies
the \emph{Hamiltonian equations}

\begin{align}
\frac{dx}{dt} & =\frac{\partial H(x,y)}{\partial y},\nonumber \\
\frac{dy}{dt} & =-\frac{\partial H(x,y)}{\partial x}.\label{eq:ham}
\end{align}
We define the map $T_{\step}(x,y)\defeq(x(\step),y(\step))$ where
the $(x(t),y(t))$ follows the Hamiltonian curve with the initial
condition $(x(0),y(0))=(x,y).$
\end{defn}

Hamiltonian Monte Carlo is a Markov chain generated by a sequence
of random Hamiltonian curves.

\begin{algorithm2e}

\caption{Hamiltonian Monte Carlo}

\SetAlgoLined

\textbf{Input:} some initial point $x^{(1)}\in\mathcal{M}$.

\For{$i=1,2,\cdots,T$}{

Sample $y$ according to $e^{-H(x^{(k)},y)}/\pi(x^{(k)})$ where $\pi(x)=\int_{\Rn}e^{-H(x,y)}dy$.

With probability $\frac{1}{2}$, set $(x^{(k+1)},y^{(k+1)})=T_{\step}(x^{(k)},y)$. 

Otherwise, $(x^{(k+1)},y^{(k+1)})=T_{-\step}(x^{(k)},y)$.

}

\textbf{Output:} $(x^{(T+1)},y^{(T+1)})$.

\end{algorithm2e}

\subsection*{Time-reversibility }
\begin{lem}[Energy Conservation]
\label{lem:ham_eng_pre}For any Hamiltonian curve $(x(t),y(t))$,
we have that
\[
\frac{d}{dt}H(x(t),y(t))=0.
\]
\end{lem}

\begin{proof}
Note that
\[
\frac{d}{dt}H(x(t),y(t))=\frac{\partial H}{\partial x}\frac{dx}{dt}+\frac{\partial H}{\partial y}\frac{dy}{dt}=\frac{\partial H}{\partial x}\frac{\partial H}{\partial y}-\frac{\partial H}{\partial y}\frac{\partial H}{\partial x}=0.
\]
\end{proof}
\begin{lem}[Measure Preservation]
\label{lem:ham_measure_pre}For any $t\geq0$, we have that
\[
\det\left(DT_{t}(x,y)\right)=1
\]
where $DT_{t}(x,y)$ is the Jacobian of the map $T_{t}$ at the point
$(x,y)$.
\end{lem}

\begin{proof}
Let $(x(t,s),y(t,s))$ be a family of Hamiltonian curves given by
$T_{t}(x+sd_{x},y+sd_{y})$. We write 
\[
u(t)=\frac{\partial}{\partial s}x(t,s)|_{s=0}\,,\,v(t)=\frac{\partial}{\partial s}y(t,s)|_{s=0}.
\]
By differentiating the Hamiltonian equations (\ref{eq:ham}) w.r.t.
$s$, we have that
\begin{align*}
\frac{du}{dt} & =\frac{\partial^{2}H(x,y)}{\partial y\partial x}u+\frac{\partial^{2}H(x,y)}{\partial y\partial y}v,\\
\frac{dv}{dt} & =-\frac{\partial^{2}H(x,y)}{\partial x\partial x}u-\frac{\partial^{2}H(x,y)}{\partial x\partial y}v,\\
(u(0),v(0)) & =(d_{x},d_{y}).
\end{align*}
This can be captured by the following matrix ODE
\begin{align*}
\frac{d\Phi}{dt} & =\left(\begin{array}{cc}
\frac{\partial^{2}H(x(t),y(t))}{\partial y\partial x} & \frac{\partial^{2}H(x(t),y(t))}{\partial y\partial y}\\
-\frac{\partial^{2}H(x(t),y(t))}{\partial x\partial x} & -\frac{\partial^{2}H(x(t),y(t))}{\partial x\partial y}
\end{array}\right)\Phi(t)\\
\Phi(0) & =I
\end{align*}
using the equation
\[
DT_{t}(x,y)\left(\begin{array}{c}
d_{x}\\
d_{y}
\end{array}\right)=\left(\begin{array}{c}
u(t)\\
v(t)
\end{array}\right)=\Phi(t)\left(\begin{array}{c}
d_{x}\\
d_{y}
\end{array}\right).
\]
Therefore, $DT_{t}(x,y)=\Phi(t)$. Next, we observe that
\[
\frac{d}{dt}\log\det\Phi(t)=\tr\left(\Phi(t)^{-1}\frac{d}{dt}\Phi(t)\right)=\tr\left(\begin{array}{cc}
\frac{\partial^{2}H(x(t),y(t))}{\partial y\partial x} & \frac{\partial^{2}H(x(t),y(t))}{\partial y\partial y}\\
-\frac{\partial^{2}H(x(t),y(t))}{\partial x\partial x} & -\frac{\partial^{2}H(x(t),y(t))}{\partial x\partial y}
\end{array}\right)=0.
\]
Hence,
\[
\det\Phi(t)=\det\Phi(0)=1.
\]
\end{proof}
Using the previous two lemmas, we can see that Hamiltonian Monte Carlo
indeed converges to the desired distribution. 
\begin{lem}[Time reversibility]
{*}\label{lem:time_reversibility}Let $p_{x}(x')$ denote the probability
density of one step of the Hamiltonian Monte Carlo starting at $x$.
We have that\textup{
\[
\pi(x)p_{x}(x')=\pi(x')p_{x'}(x)
\]
for almost everywhere in $x$ and $x'$ where $\pi(x)=\int_{\Rn}e^{-H(x,y)}dy$.}
\end{lem}

\begin{proof}
Fix $x$ and $x'$. Let $F_{\step}^{x}(y)$ be the $x$ component
of $T_{\step}(x,y)$. Let $V_{+}=\{y:\text{ }F_{\step}^{x}(y)=x'\}$
and $V_{-}=\{y:\text{ }F_{-\step}^{x}(y)=x')\}$. Then, 
\[
\pi(x)p_{x}(x')=\frac{1}{2}\int_{y\in V_{+}}\frac{e^{-H(x,y)}}{\left|\det\left(DF_{\step}^{x}(y)\right)\right|}+\frac{1}{2}\int_{y\in V_{-}}\frac{e^{-H(x,y)}}{\left|\det\left(DF_{-\step}^{x}(y)\right)\right|}.
\]
We note that this formula assumed that $DF_{\step}^{x}$ is invertible.
Sard's theorem showed that $F_{\step}^{x}(N)$ is measure $0$ where
$N\defeq\{y:\,DF_{s}^{x}(y)\text{ is not invertible}\}$. Therefore,
the formula is correct except for a measure zero subset.

By reversing time for the Hamiltonian curve, we have that for the
same $V_{\pm}$, 
\begin{equation}
\pi(x')p_{x'}(x)=\frac{1}{2}\int_{y\in V_{+}}\frac{e^{-H(x',y')}}{\left|\det\left(DF_{-\step}^{x'}(y')\right)\right|}+\frac{1}{2}\int_{y\in V_{-}}\frac{e^{-H(x',y')}}{\left|\det\left(DF_{\step}^{x'}(y')\right)\right|}\label{eq:ham_transit}
\end{equation}
where $y'$ denotes the $y$ component of $T_{\step}(x,y)$ and $T_{-\step}(x,y)$
in the first and second sum respectively.

We compare the first terms in both equations. Let $DT_{\step}(x,y)=\left(\begin{array}{cc}
A & B\\
C & D
\end{array}\right)$. Since $T_{\step}\circ T_{-\step}=I$ and $T_{\step}(x,y)=(x',y')$,
the inverse function theorem shows that $DT_{-\step}(x',y')$ is the
inverse map of $DT_{\step}(x,y)$. Hence, we have that
\[
DT_{-\step}(x',y')=\left(\begin{array}{cc}
A & B\\
C & D
\end{array}\right)^{-1}=\left(\begin{array}{cc}
\cdots & -A^{-1}B(D-CA^{-1}B)^{-1}\\
\cdots & \cdots
\end{array}\right).
\]
Therefore, we have that $F_{\step}^{x}(y)=B$ and $F_{-\step}^{x'}(y')=-A^{-1}B(D-CA^{-1}B)^{-1}$.
Hence, we have that
\begin{align*}
\left|\det\left(DF_{-\step}^{x'}(y')\right)\right| & =\left|\det A^{-1}\det B\det\left(D-CA^{-1}B\right)^{-1}\right|=\frac{\left|\det B\right|}{\left|\det\left(\begin{array}{cc}
A & B\\
C & D
\end{array}\right)\right|}.
\end{align*}
Using that $\det\left(DT_{t}(x,y)\right)=\det\left(\begin{array}{cc}
A & B\\
C & D
\end{array}\right)=1$ (Lemma \ref{lem:ham_measure_pre}), we have that
\[
\left|\det\left(DF_{-\step}^{x'}(y')\right)\right|=\left|\det\left(DF_{\step}^{x}(y)\right)\right|.
\]
Hence, we have that
\begin{align*}
\frac{1}{2}\int_{y\in V_{+}}\frac{e^{-H(x,y)}}{\left|\det\left(DF_{\step}^{x}(y)\right)\right|} & =\frac{1}{2}\int_{y\in V_{+}}\frac{e^{-H(x,y)}}{\left|\det\left(DF_{-\step}^{x'}(y')\right)\right|}\\
 & =\frac{1}{2}\int_{y\in V_{+}}\frac{e^{-H(x',y')}}{\left|\det\left(DF_{-\step}^{x'}(y')\right)\right|}
\end{align*}
where we used that $e^{-H(x,y)}=e^{-H(x',y')}$ (Lemma \ref{lem:ham_eng_pre})
at the end. 

For the second term in (\ref{eq:ham_transit}), by the same calculation,
we have that
\[
\frac{1}{2}\int_{y\in V_{-}}\frac{e^{-H(x,y)}}{\left|\det\left(DF_{-\step}^{x}(y)\right)\right|}=\frac{1}{2}\int_{y\in V_{+}}\frac{e^{-H(x',y')}}{\left|\det\left(DF_{\step}^{x'}(y')\right)\right|}
\]
\end{proof}

\subsection*{Convergence}

First we consider the convergence in the case when $H(x,v)=f(x)+\frac{1}{2}\norm v^{2}$
for a strongly convex function $f$. So the marginal of the stationary
distribution along $x$ is proportional to $e^{-f}$. The idea here
is coupling (as we did for Langevin dynamics). We consider two separate
processes $x$ and $y$, with their next step directions chosen to
be identical. The key lemma is that with this setting the squared
distance decreases up to a certain time that depends on the condition
number.
